### 2021-09-30 GLOBI_bee_update_clean.Rmd
Adds script to pull bee / plant data from GLOBI. Also cleans data such that bees are source and plants are targets.

---
title:          Data Cleaning
output:         html_document
participants:   N. Bachelder, A. Chen, Z. Fang, M. Rapaport, S. Solomon
date:           Feb 18, 2021
source:         https://www.globalbioticinteractions.org/about
---

## Read data
```{r}
library(data.table)

df <- fread( 'https://zenodo.org/record/4460654/files/interactions.csv.gz')
```

# Filtering
```{r}
# List of total seven of bee families
b_families = c('Andrenidae','Apidae','Colletidae','Halictidae','Megachilidae','Melittidae','Stenotritidae')

# Bee as target, Plant as source
bee_source_df <- df %>% filter(sourceTaxonFamilyName %in% b_families, targetTaxonKingdomName == 'Plantae')

# Plant as source, Bee as target
plant_source_df = df %>% filter(sourceTaxonKingdomName == 'Plantae', targetTaxonFamilyName %in% b_families)

# Flip target/source for above df

colnames(plant_source_df) <- as.character(lapply(colnames(plant_source_df), textclean::swap, pattern1 = 'source', pattern2 = 'target'))

plant_source_df <- plant_source_df %>% rename(sourceCitation = targetCitation, sourceNamespace = targetNamespace, 
                                              sourceArchiveURI = targetArchiveURI, sourceDOI = targetDOI,
                                              sourceLastSeenAtUnixEpoch = targetLastSeenAtUnixEpoch)
```

## Combine
```{r}
# Combine without dropping duplicates
# The filtering method guarantees two disjoint data frames
interactions_dup = rbind(bee_source_df, plant_source_df)

# Combine and drop duplicates
interactions = rbind(bee_source_df, plant_source_df) %>% unique()
```

## Result
```{r}
interactions
```

# Export to csv.gz file
```{r}
write.csv(interactions, "generated_data/interactions.csv")
write.csv(interactions_dup, "generated_data/interactions_dup.csv")
```

### 2021-10-01 GLOBI-Fowler Data Prep.Rmd
Adds GLOBI-Fowler Modeling Prep. Also adds some GLOBI data exploration from last quarter (specifically on andrenidae bees)

---
title: "BeeModeling"
author: "Nick R. Bachelder"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(dplyr)
library(data.table)
```

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('interactions_dup.csv') %>% select(-X)

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}
```

```{r, echo = False}
beeCiteW
```

Define beeCites with varying criteria

```{r, fig.width=5,fig.height=2}
beeCiteNoFilter <- interactions_b %>% select(sourceTaxonSpeciesName, targetTaxonFamilyName) %>% dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% group_by(interestBee, interact) %>% count() %>% ungroup() %>% filter(interestBee != "", interact != "")

TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(freq)) %>% ungroup()
TotalCitations <- sum(TotalCitations$n_Bee)

beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = freq, n_Bee = sum(n)) %>% ungroup()
beeCiteNoFilter <- beeCiteNoFilter %>% summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()

beeIntDegree <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(freq))

beeIntDegreeC10 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(freq)) ## min of 10 citations to contribute to degree

beeIntDegreeC5 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(freq)) ## min of 5 citations to contribute to degree

beeIntDegreeW <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeW$numcitations <- beeCiteNoFilter$n
beeIntDegreeW$n_Bee <- beeCiteNoFilter$n_Bee
beeIntDegreeW$n <- (beeIntDegreeW$numcitations / beeIntDegreeW$n_Bee)^2
beeIntDegreeW <- beeIntDegreeW %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## weighted degree

beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type="inner") %>% filter(n > 5)

beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee", type="inner") %>% filter(n > 5)

beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee", type="inner") %>% filter(n > 5)
beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee", type="inner") %>% filter(n > 5)
```

## Include Fowler Data

```{r}
western_page <- read_html("https://jarrodfowler.com/pollen_specialist.html")
eastern_page <- read_html("https://jarrodfowler.com/specialist_bees.html")
central_page <- read_html("https://jarrodfowler.com/bees_pollen.html")

western_table <- western_page %>% html_table(fill=TRUE)
eastern_table<- eastern_page %>% html_table(fill=TRUE)
central_table<- central_page %>% html_table(fill=TRUE)

western_table <- as.data.frame(western_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
eastern_table <- as.data.frame(eastern_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
central_table <- as.data.frame(central_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)

western_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", western_table$Genus..Subgenus..species))
eastern_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", eastern_table$Genus..Subgenus..species))
central_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", central_table$Genus..Subgenus..species))

table <- rbind(western_table, eastern_table, central_table)
table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", table$Genus..Subgenus..species))

western_table$west_specialist <- 1
western_test <- western_table %>% dplyr::select(interestBee, west_specialist) %>% unique()

eastern_table$east_specialist <- 1
eastern_test <- eastern_table %>% dplyr::select(interestBee, east_specialist) %>% unique()

central_table$cent_specialist <- 1
central_test <- central_table %>% dplyr::select(interestBee, cent_specialist) %>% unique()

tablejoining <- merge(x = western_test, y = eastern_test, by = "interestBee", all = TRUE)
table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)
table_test[is.na(table_test)] <- 0
```

```{r}
plotdf <- merge(beeCite, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

fit = glm(specialist ~ n_Bee + degree, data = plotdf, family = 'binomial')

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 40))+ geom_jitter()+ scale_color_discrete(name = "Specialist") + geom_smooth()

write.csv(plotdf, 'BeeCiteRaw')
```
```{r}
plotdf <- merge(beeCiteC5, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist)), text = paste("Bee:", interestBee)) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity (Minimum 5 Citations) and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 15))+ geom_jitter() + scale_color_discrete(name = "Specialist")+ geom_smooth(se = FALSE)

write.csv(plotdf, 'BeeCiteC5')
```

```{r}
plotdf <- merge(beeCiteW, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Function of Degree')) + ggtitle(paste('Function of Degree and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 1)) + geom_smooth(se = FALSE)+ scale_color_discrete(name = "Specialist")

write.csv(plotdf, 'BeeCiteW')
```

### 2021-10-01 exploration/Andrenidae Data Exloration.Rmd
folder organization

---
title: "Data Heatmaps"
author: "Nick R. Bachelder"
date: "2/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(igraph)
library(ggplot2)
library(tidyverse)
library(viridis)
library(plyr)

options(max.print=1000000)
interactions_b <- read.csv('interactions_dup.csv') %>% select(-X)

interactions_b[c(1,2,3,4,5),] 
```

DO FOR SPECIFICALLY "adrenidae"
```{r}
for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

interactions_andrenidae <- interactions_b %>% filter(sourceTaxonFamilyName == 'andrenidae')
```

Heatmap for Recorded Ineractions of Family of Bee and Plantae Families.
```{r}
beeCite2 <- interactions_andrenidae %>% dplyr::select(c(sourceTaxonFamilyName, targetTaxonFamilyName)) %>%
  dplyr::rename(interestBee = sourceTaxonFamilyName, interact = targetTaxonFamilyName) %>%
  group_by(interestBee, interact) %>% count() %>% filter(freq >= 300) %>% ungroup() %>% complete(interestBee, interact, fill = list(freq = 0))

beeCite2 %>% ggplot( aes(x=interact, y= freq)) +
    geom_segment( aes(x=interact ,xend=interact, y=0, yend= freq), color="black") +
    geom_point(size=4, color="#69b3a2") +
    coord_flip() + xlab("Plantae Genus") + ylab('Number of Interactions') + ggtitle('Ineractions of Andrenidae Bee Family and Various Plantae Families')

beeCite2

```

Heatmap for Recorded Ineractions of Genus of Bee and Plantae Genus
```{r}
beeCite2 <- interactions_andrenidae %>% dplyr::select(c(sourceTaxonGenusName, targetTaxonGenusName)) %>% dplyr::rename(interestBee = sourceTaxonGenusName, interact = targetTaxonGenusName) %>%
  group_by(interestBee, interact) %>% count() %>% filter(freq >=400) %>% ungroup() %>% complete(interestBee, interact, fill = list(freq = 0))

ggplot(beeCite2, aes(interestBee, interact, fill= freq)) +
  geom_tile(color = 'black') + xlab(paste('Genus of Bee')) + labs(fill = "Number of Citations") + ylab(paste('Plant Genus Bees Interact With')) + 
  ggtitle(paste('Recorded Ineractions of Genus of Andrenidae Bee and Plantae Genus')) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Heatmap for Recorded Ineractions of Family of Bee and Plantae Genus
```{r}
beeCite2 <- interactions_andrenidae %>% dplyr::select(c(sourceTaxonFamilyName, targetTaxonGenusName)) %>% dplyr::rename(interestBee = sourceTaxonFamilyName, interact = targetTaxonGenusName) %>%
  group_by(interestBee, interact) %>% count() %>% filter(freq>=500) %>% ungroup() %>% complete(interestBee, interact, fill = list(freq = 0))

beeCite2 %>% ggplot( aes(x=interact, y=freq)) +
    geom_segment( aes(x=interact ,xend=interact, y=0, yend= freq), color="black") +
    geom_point(size=4, color="#69b3a2") +
    coord_flip() + xlab("Plantae Genus") + ylab('Number of Interactions') + ggtitle('Ineractions of Andrenidae Bee Family and Plantae Genus')

```

Heatmap for Recorded Ineractions of Genus of Bee and Plantae Family
```{r}
beeCite2 <- interactions_andrenidae %>% dplyr::select(c(sourceTaxonGenusName, targetTaxonFamilyName)) %>% dplyr::rename(interestBee = sourceTaxonGenusName, interact = targetTaxonFamilyName) %>%
  group_by(interestBee, interact) %>% count() %>% filter(freq>=600) %>% ungroup() %>% complete(interestBee, interact, fill = list(freq = 0))

ggplot(beeCite2, aes(interestBee, interact, fill= freq)) +
  geom_tile(color = 'black') + xlab(paste('Genus of Bee')) + labs(fill = "Number of Citations") + ylab(paste('Plant Family Bees Interact With')) + 
  ggtitle(paste('Recorded Ineractions of Genus of Andrenidae Bee and Plantae Families')) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

#### WEEK 9 : FIX Heatmaps

I determined that the heatmaps are slightly flawed in the sense that, when the filter is applied, it wipes all the relationships below the filter. Then, subsequently, it refills these relationships with 0 if the bee level has another plant relationship that is above the threshold. Therefore, there are values of bee relationships that are listed as 0, but should be somewhere between 0 and the filter level. Here is my solution:

Here is a flawed heatmap for family to family. Many of these 0s are slightly higher than 0.

```{r}
ggplot(beeCite2, aes(interestBee, interact, fill= freq)) +
  geom_tile(color = 'black') + xlab(paste('Genus of Bee')) + labs(fill = "Number of Citations") + ylab(paste('Plant Family Bees Interact With')) + 
  ggtitle(paste('Recorded Ineractions of Family of Bee and Plantae Families')) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ scale_fill_gradient(
    low = "white",
    high = "blue",
    guide = "colourbar"
  ) + geom_text(aes(label = freq))
```

### 2021-10-01 MS_data_modeling/MS_data_explore.Rmd
Pulls MS data in. Compares fowler, GLOBI, and MS dat

---
title: "MS_data_clean"
author: "Nick R. Bachelder"
date: "10/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lifecycle)
library(rvest)
library(dplyr)
library(plotly)
library(dplyr)
library(data.table)
library(stringr)
library(readxl)

MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_09-26-21_ALR_seltmann-colab.xlsx', sheet = 2)

globee <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)

fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
   mutate(interestBee = str_to_title(interestBee)) %>%
  select(interestBee, west_specialist, east_specialist, cent_specialist, specialist)
```

```{r}
MS_data_format <- MS_data %>% mutate(interestBee = str_to_title(paste(Genus, sub(".*? ", "", Species)))) %>% 
  rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth') %>%
  select(interestBee, degreeMS, specializationMS) %>%
  filter(!is.na(degreeMS)) %>% arrange(interestBee)

MS_data_format
```

```{r}
globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE = degree) %>% 
  select(interestBee, degreeGLOBEE) %>% unique() %>%
  arrange(interestBee)

globee_format
```

```{r}
compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), all = T)

compare_df <- compare_df %>% merge(fowler %>% rename(specialistFOWLER = specialist) %>% 
                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% 
  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(degreeMS), 0, 1), 
         existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
         existsALL = existsGLOBEE*existsMS*existsFOWLER)

compare_df %>% filter(existsALL == 1)
```

```{r}

```

```{r}

```

```{r}

```

### 2021-10-08 MS_data_modeling/MS_data_explore.Rmd
Adds specialization comparisons, bee sources table.

@@
 library(dplyr)
+library(tidyr)
 library(plotly)
@@
 library(readxl)
+library(kableExtra)
 
@@
 globee <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
+  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+
+globee_5cut <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteC5') %>%
+  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+
+globee_T <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteW') %>%
   select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
@@
   filter(!is.na(degreeMS)) %>% arrange(interestBee)
-
-MS_data_format
 ```
@@
 
-globee_format
+globee_5cut_format <- globee_5cut %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE_C5 = degree) %>% 
+  select(interestBee, degreeGLOBEE_C5) %>% unique() %>%
+  arrange(interestBee)
+
+globee_T_format <- globee_T %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE_T = degree) %>% 
+  select(interestBee, degreeGLOBEE_T) %>% unique() %>%
+  arrange(interestBee)
 ```
@@
 compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), all = T)  
+compare_df <- merge(compare_df, globee_5cut_format, by = c('interestBee'), all = T) 
+compare_df <- merge(compare_df, globee_T_format, by = c('interestBee'), all = T)
 
 compare_df <- compare_df %>% merge(fowler %>% rename(specialistFOWLER = specialist) %>% 
-                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% 
+                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% unique() %>%
   mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(degreeMS), 0, 1), 
@@
          existsALL = existsGLOBEE*existsMS*existsFOWLER)
+```
 
-compare_df %>% filter(existsALL == 1)
+### Add individual citation cuts in addition to raw GLOBI
+### Figure out what we want to define as 'specialist' MS data
+
+```{r}
+only_globi <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 0, existsFOWLER == 0) %>% nrow()
+globi_fowler <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 0, existsFOWLER == 1) %>% nrow()
+globi_MS <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 1, existsFOWLER == 0) %>% nrow()
+MS_fowler <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 1, existsFOWLER == 1) %>% nrow()
+fowler_only <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 0, existsFOWLER == 1) %>% nrow()
+MS_only <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 1, existsFOWLER == 0) %>% nrow()
+all_sources <- compare_df %>% filter(existsALL == 1) %>% nrow()
+
+counts <- c(only_globi,fowler_only,MS_only, globi_MS, globi_fowler, MS_fowler, all_sources, nrow(compare_df))
+data_sources <- c('Only in GLOBI', 'Only in Fowler', 'Only in MS', 'GLOBI and MS', 'GLOBI and Fowler', 'MS and Fowler', 'All Sources', 'Total')
+
+data.frame(data_sources, counts) %>% mutate('Percent of Data' = counts / nrow(compare_df)) %>% 
+  rename('Data Source' = data_sources, 'Unique Genus Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
+  kable_styling()
 ```
+
+```{r}
+compare_df %>% filter(existsALL == 1) %>% group_by(specializationMS) %>% summarize('Mean GLOBI Degree'  = mean(degreeGLOBEE),
+                                                                                   'GLOBI Degree IQR' = paste('[', round(quantile(degreeGLOBEE, 0.25),
+                                                                                                                         2), 
+                                                                                                                  ',', round(quantile(degreeGLOBEE, 0.75), 2),
+                                                                                                            ']'),
+                                                                                   'Mean GLOBI C5 Degree' = mean(degreeGLOBEE_C5),
+                                                                                   'GLOBI C5 Degree IQR' = paste('[', round(quantile(degreeGLOBEE_C5, 0.25), 2),
+                                                                                                                  ',', round(quantile(degreeGLOBEE_C5, 0.75), 2),
+                                                                                                            ']'),
+                                                                                   'Mean GLOBI T Degree' = mean(degreeGLOBEE_T),
+                                                                                   'GLOBI T Degree IQR' = paste('[', round(quantile(degreeGLOBEE_T, 0.25), 2),
+                                                                                                                  ',', round(quantile(degreeGLOBEE_T, 0.75), 2),
+                                                                                                            ']'),
+                                                                                   'Mean MS Degree' = mean(degreeMS),
+                                                                                   'MS Degree IQR' = paste('[', round(quantile(degreeMS, 0.25), 2), 
+                                                                                                                  ',', round(quantile(degreeMS, 0.75), 2),
+                                                                                                            ']'),
+                                                                                   'Percent Fowler Specialist' = mean(specialistFOWLER)
+) %>% kbl(caption = 'Plant Family Degrees by MS Specialization') %>%
+  kable_styling()
+```
+
+```{r}
+# Bees fowler specialists but not MS specialists
+
+compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
+  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5)
+```
+
+```{r}
+# Bees NOT fowler specialists but MS specialists
+
+compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
+  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5)
+```
+
+## Plan: want to add pollen hosts so I can filter on pollen family. Discuss how to parse the authors out of appropriate MS column.
 

### 2021-10-15 MS_data_modeling/MS_data_explore.Rmd
Adds MS sources and pollen hosts. Cleans columns by removing author abbreviations.

@@
 MS_data_format <- MS_data %>% mutate(interestBee = str_to_title(paste(Genus, sub(".*? ", "", Species)))) %>% 
-  rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth') %>%
-  select(interestBee, degreeMS, specializationMS) %>%
-  filter(!is.na(degreeMS)) %>% arrange(interestBee)
+  rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth', pollenHosts = 'Pollen hosts') %>%
+  select(interestBee, pollenHosts, degreeMS, specializationMS, References) %>%
+  filter(!is.na(degreeMS)) %>% arrange(interestBee) %>% mutate(pollenHosts = str_to_title(gsub("[A-Za-z]*[^A-Za-z/, ]\\S*\\s*", "", pollenHosts)))
+
+MS_data_format
 ```
@@
 compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5)
+  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE)
 ```
@@
 compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5)
+  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE) %>% filter(degreeGLOBEE_C5 > 5)
 ```
 
-## Plan: want to add pollen hosts so I can filter on pollen family. Discuss how to parse the authors out of appropriate MS column.
+## Plan: want to add pollen hosts so I can filter on pollen family. Discuss how to parse the authors out of appropriate MS column. ## DONE
+
+## Cleaning up source and pollen columns (removin author abbreviation)   ### DONE
+
+## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa)
+
+## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables)
 

### 2021-10-15 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep.Rmd
Refilters fowler so its not filtered only to GLOBI bees. Adds the pollen hosts as listed by GLOBI (to be compared to pollen hosts in MS data). Reruns tables to reflect changes.

--- 2021-10-01 GLOBI-Fowler Data Prep.Rmd
@@
 options(max.print=1000000)
-interactions_b <- read.csv('interactions_dup.csv') %>% select(-X)
+interactions_b <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv') %>% select(-X)
 
@@
 
-```{r, echo = False}
-beeCiteW
-```
-
 Define beeCites with varying criteria
@@
 
-TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(freq)) %>% ungroup()
+TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(n)) %>% ungroup()
 TotalCitations <- sum(TotalCitations$n_Bee)
 
-beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = freq, n_Bee = sum(n)) %>% ungroup()
+beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
 beeCiteNoFilter <- beeCiteNoFilter %>% summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()
@@
 beeIntDegree <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
-beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(freq))
+beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(n))
 
@@
 beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
-beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(freq)) ## min of 10 citations to contribute to degree
+beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 10 citations to contribute to degree
 
@@
 beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
-beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(freq)) ## min of 5 citations to contribute to degree
+beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 5 citations to contribute to degree
 
@@
 
-beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type="inner") %>% filter(n > 5)
+beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type = 'inner')
 
-beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee", type="inner") %>% filter(n > 5)
+beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee",  type = 'inner')
 
-beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee", type="inner") %>% filter(n > 5)
-beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee", type="inner") %>% filter(n > 5)
+beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee",  type = 'inner')
+
+beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee",  type = 'inner')
+
+beeCiteW
 ```
@@
 table_test[is.na(table_test)] <- 0
+
+table_test$specialist <- table_test$west_specialist + table_test$east_specialist + table_test$cent_specialist
+table_test$specialist <- as.integer(table_test$specialist > 0)
+table_test$specialist[is.na(table_test$specialist)] <- 0
+
+write.csv(table_test, 'fowler')
 ```
@@
 ```
+
 ```{r}

### 2021-10-15 MS_data_modeling/MS_data_explore.Rmd
Refilters fowler so its not filtered only to GLOBI bees. Adds the pollen hosts as listed by GLOBI (to be compared to pollen hosts in MS data). Reruns tables to reflect changes.

@@
 
-fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
+fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/fowler') %>%
    mutate(interestBee = str_to_title(interestBee)) %>%
@@
 ```{r}
+pollenHostGLOBEE <- aggregate(interact~interestBee, globee_5cut, FUN=toString) %>% mutate(interestBee = str_to_title(interestBee)) %>% 
+  rename('pollenHostGLOBEE' = interact)
+## this is plant interactions for each bee genus according to globi
+
 globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE = degree) %>% 
   select(interestBee, degreeGLOBEE) %>% unique() %>%
-  arrange(interestBee)
+  arrange(interestBee) %>% merge(pollenHostGLOBEE, by = 'interestBee')
 
@@
          existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
-         existsALL = existsGLOBEE*existsMS*existsFOWLER)
+         existsALL = existsGLOBEE*existsMS*existsFOWLER,
+         specialistFOWLER = ifelse(is.na(specialistFOWLER), 0 , specialistFOWLER),
+         pollenHostGLOBEE = str_to_title(pollenHostGLOBEE)) %>% rename(pollenHostsMS = pollenHosts,
+                                                                                              ReferencesMS = References) %>%
+  select(interestBee, specializationMS, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, specialistFOWLER, 
+         existsGLOBEE, existsMS, existsFOWLER, existsALL, pollenHostsMS, pollenHostGLOBEE, ReferencesMS)
+
+compare_df %>% filter(existsALL == 1)
 ```
@@
 compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE)
+  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHosts) %>% filter(degreeGLOBEE_C5 > 3)
 ```
@@
 
-## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa)
+## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa) ## WIP
 
-## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables)
+## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables). ## DONE: Turns out that we didnt cut globi bees, but we instead cut FOWLER bees.
+
+### Also: Added the target pollen hosts from globi so we can compare to pollen hosts from MS
 
 ```{r}
-
+compare_df
 ```
@@
 ```
-
-```{r}
-
-```

### 2021-10-15 MS_data_modeling/MS_data_explore.Rmd
Defines odd bees (both directions)

@@
          existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
-         existsALL = existsGLOBEE*existsMS*existsFOWLER,
-         specialistFOWLER = ifelse(is.na(specialistFOWLER), 0 , specialistFOWLER),
-         pollenHostGLOBEE = str_to_title(pollenHostGLOBEE)) %>% rename(pollenHostsMS = pollenHosts,
-                                                                                              ReferencesMS = References) %>%
+         existsALL = existsGLOBEE*existsMS,
+         pollenHostGLOBEE = str_to_title(pollenHostGLOBEE)) %>%
+  mutate(specialistFOWLER = ifelse(is.na(specialistFOWLER), 0 , specialistFOWLER))%>% 
+  rename(pollenHostsMS = pollenHosts,ReferencesMS = References) %>%
   select(interestBee, specializationMS, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, specialistFOWLER, 
-         existsGLOBEE, existsMS, existsFOWLER, existsALL, pollenHostsMS, pollenHostGLOBEE, ReferencesMS)
+         existsGLOBEE, existsMS, existsFOWLER, existsALL, pollenHostsMS, pollenHostGLOBEE, ReferencesMS, degreeMS)
 
-compare_df %>% filter(existsALL == 1)
+compare_df %>% filter(existsALL == 0)
 ```
@@
                                                                                                             ']'),
-                                                                                   'Mean MS Degree' = mean(degreeMS),
-                                                                                   'MS Degree IQR' = paste('[', round(quantile(degreeMS, 0.25), 2), 
-                                                                                                                  ',', round(quantile(degreeMS, 0.75), 2),
-                                                                                                            ']'),
                                                                                    'Percent Fowler Specialist' = mean(specialistFOWLER)
@@
 compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHosts) %>% filter(degreeGLOBEE_C5 > 3)
+  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE < 3)
 ```
@@
 compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE) %>% filter(degreeGLOBEE_C5 > 5)
+  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 > 5)
 ```
@@
 ```{r}
-compare_df
+
 ```
@@
 ```
+
+```{r}
+
+```

### 2021-11-29 MS_data_modeling/MS_data_explore.Rmd
modeling updates, family level data transofmation on MS data via cat of life

--- 2021-10-15 MS_data_modeling/MS_data_explore.Rmd
@@
 library(kableExtra)
+library(rcol)
+library(plyr)
 
-MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_09-26-21_ALR_seltmann-colab.xlsx', sheet = 2)
+MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_10-26-21_ALR seltmann-colab.xlsx', sheet = 2)
 
 globee <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
-  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
 globee_5cut <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteC5') %>%
-  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
 globee_T <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteW') %>%
-  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
-fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
+fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/fowler') %>%
    mutate(interestBee = str_to_title(interestBee)) %>%
-  select(interestBee, west_specialist, east_specialist, cent_specialist, specialist)
+  dplyr::select(interestBee, west_specialist, east_specialist, cent_specialist, specialist) %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
 ```
@@
 MS_data_format <- MS_data %>% mutate(interestBee = str_to_title(paste(Genus, sub(".*? ", "", Species)))) %>% 
-  rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth', pollenHosts = 'Pollen hosts') %>%
-  select(interestBee, pollenHosts, degreeMS, specializationMS, References) %>%
-  filter(!is.na(degreeMS)) %>% arrange(interestBee) %>% mutate(pollenHosts = str_to_title(gsub("[A-Za-z]*[^A-Za-z/, ]\\S*\\s*", "", pollenHosts)))
+  dplyr::rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth', pollenHosts = 'Pollen hosts') %>%
+  dplyr::select(interestBee, pollenHosts, degreeMS, specializationMS, References) %>%
+  filter(!is.na(specializationMS)) %>% arrange(interestBee) %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
 
-MS_data_format
+MS_data_format %>% filter(!(is.na(specializationMS)), is.na(pollenHosts)) ### there
+```
+
+### create family level column by pulling
+
+```{r}
+cp_nu_search(q='Sesamum', status = "accepted")$result$classification
 ```
@@
 ```{r}
-globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE = degree) %>% 
-  select(interestBee, degreeGLOBEE) %>% unique() %>%
-  arrange(interestBee)
+MS_data_format$pollenHosts <-  gsub("\\?.*", "", MS_data_format$pollenHosts)
+MS_data_format$pollenHosts <-  gsub("\\..*", "", MS_data_format$pollenHosts)
 
-globee_5cut_format <- globee_5cut %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE_C5 = degree) %>% 
-  select(interestBee, degreeGLOBEE_C5) %>% unique() %>%
-  arrange(interestBee)
+ms_pollen_hosts = na.omit(unique(
 
-globee_T_format <- globee_T %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE_T = degree) %>% 
-  select(interestBee, degreeGLOBEE_T) %>% unique() %>%
-  arrange(interestBee)
+  sub("\\/.*", "", word(unlist(str_split(unique(MS_data_format$pollenHosts), pattern = ', ')), 1))
+
+  ))
+
+## pulls all unique maximum level taxonomic names from MS pollen hosts
+ms_pollen_hosts <- ms_pollen_hosts[!(ms_pollen_hosts == '')]
+
+ms_pollen_hosts
+```
+
+```{r}
+library(rcol)
+library(pbapply)
+
+find_family <- function(name) {
+  try <- try(cp_nu_search(q=name, status = "accepted"), silent=TRUE)
+  if ("try-error" %in% class(try)) {
+    label <- 'Unknown'
+  }
+  try <- try(str_to_title(tolower(
+    (rbind( data.frame(try$result$classification[1]),
+        data.frame(try$result$classification[2]),
+       data.frame(try$result$classification[3])) %>% filter(rank == 'family'))[1,]$label
+    )), silent=TRUE)
+  if ("try-error" %in% class(try)) {
+    label <- 'Unknown'
+  }
+  else {
+    label <- try
+  }
+  return(label)
+}
+
+familys <- pblapply(ms_pollen_hosts, find_family)
+```
+
+```{r}
+familys <- unlist(familys)
+family_key <- data.frame(cbind(ms_pollen_hosts, familys)) %>% `colnames<-`(c('max_taxons', 'pollen_host_familys'))
+
+family_key
+```
+
+## Make the below two entries have 0 rows with Avery's help
+```{r}
+family_key %>% filter(is.na(pollen_host_familys))
+```
+
+```{r}
+family_key %>% filter(pollen_host_familys == 'Unknown', grepl( ',', max_taxons, fixed = TRUE))
+```
+
+## make sure the below has no family name synonyms. Take out the unknownand NA values until we fix them with avery.
+
+```{r}
+pollen_family_MS <- MS_data_format %>% select(interestBee, pollenHosts) %>% separate_rows(pollenHosts, sep=', ') %>% mutate(max_taxons = word(pollenHosts, 1)  ) %>% left_join(family_key, by = 'max_taxons') %>% select(interestBee, pollen_host_familys) %>%
+  group_by(interestBee) %>% 
+  dplyr::summarise(pollen_host_familys = paste(unique(as.vector(pollen_host_familys)), collapse = ', '))
+
+MS_data_format <- MS_data_format %>% left_join(pollen_family_MS, by = 'interestBee') 
+```
+
+```{r}
+avery_send <- MS_data_format %>% select(interestBee, pollenHosts, pollen_host_familys) %>% mutate(ifelse(pollen_host_familys == 'NA', NA, pollen_host_familys))
+
+write.csv(avery_send, "/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Family_Level_Transform.csv")
+
+avery_send
+```
+
+### Clean GLOBEE data
+
+```{r}
+pollenHostGLOBEE <- aggregate(interact~interestBee, globee_5cut, FUN=toString) %>% mutate(interestBee = str_to_title(interestBee)) %>% 
+  dplyr::rename('pollenHostGLOBEE' = interact) %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+## this is plant interactions for each bee genus according to globi
+
+globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE = degree) %>% 
+  dplyr::select(interestBee, degreeGLOBEE, n_Bee) %>% unique() %>%
+  arrange(interestBee) %>% merge(pollenHostGLOBEE, by = 'interestBee') %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+globee_5cut_format <- globee_5cut %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_C5 = degree) %>% 
+  dplyr::select(interestBee, degreeGLOBEE_C5) %>% unique() %>%
+  arrange(interestBee)%>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+globee_T_format <- globee_T %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_T = degree) %>% 
+  dplyr::select(interestBee, degreeGLOBEE_T) %>% unique() %>%
+  arrange(interestBee)%>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
 ```
@@
 
-compare_df <- compare_df %>% merge(fowler %>% rename(specialistFOWLER = specialist) %>% 
-                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% unique() %>%
-  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(degreeMS), 0, 1), 
+compare_df <- compare_df %>% merge(fowler %>% dplyr::rename(specialistFOWLER = specialist) %>% 
+                                     dplyr::select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% unique() %>%
+  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(specializationMS), 0, 1), 
          existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
-         existsALL = existsGLOBEE*existsMS*existsFOWLER)
+         existsALL = existsGLOBEE*existsMS,
+         pollenHostGLOBEE = str_to_title(pollenHostGLOBEE)) %>%
+  mutate(specialistFOWLER = ifelse(is.na(specialistFOWLER), 0 , specialistFOWLER))%>% 
+  dplyr::rename(pollenHostsMS = pollenHosts ,pollenHostsFamilysMS = pollen_host_familys,ReferencesMS = References, nBeeGLOBEE = n_Bee) %>%
+  dplyr::select(interestBee, specializationMS, nBeeGLOBEE, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, specialistFOWLER, 
+         existsGLOBEE, existsMS, existsFOWLER, existsALL, pollenHostsMS, pollenHostsFamilysMS, pollenHostGLOBEE, ReferencesMS, degreeMS) %>%
+  rowwise() %>%
+  dplyr::mutate(pollenHostsFamilysMS = ifelse(pollenHostsFamilysMS == 'NA', NA, pollenHostsFamilysMS) ) %>%
+  dplyr::mutate(FamilyDegreeMS = ifelse(!is.na(pollenHostsFamilysMS), length(strsplit(pollenHostsFamilysMS, split = ', ')[[1]]), 0  ) )
+
+compare_df
 ```
@@
 data.frame(data_sources, counts) %>% mutate('Percent of Data' = counts / nrow(compare_df)) %>% 
-  rename('Data Source' = data_sources, 'Unique Genus Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
+  dplyr::rename('Data Source' = data_sources, 'Unique Species Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
   kable_styling()
@@
 ```{r}
-compare_df %>% filter(existsALL == 1) %>% group_by(specializationMS) %>% summarize('Mean GLOBI Degree'  = mean(degreeGLOBEE),
+compare_df %>% filter(existsALL == 1) %>% group_by(specializationMS) %>% dplyr::summarize('Mean GLOBI Degree'  = mean(degreeGLOBEE),
                                                                                    'GLOBI Degree IQR' = paste('[', round(quantile(degreeGLOBEE, 0.25),
@@
 
-compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE)
+odd_generalists <- compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
+  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsFamilysMS, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 == 1)
+
+odd_generalists
+
+write.csv(odd_generalists, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/odd_polylectic.csv')
 ```
@@
 
-compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE) %>% filter(degreeGLOBEE_C5 > 5)
+odd_specialists <- compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
+  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsMS, pollenHostsFamilysMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 > 3)
+
+odd_specialists
+
+write.csv(odd_specialists, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/odd_specialists.csv')
 ```
 
-## Plan: want to add pollen hosts so I can filter on pollen family. Discuss how to parse the authors out of appropriate MS column. ## DONE
+```{r}
+# Bees NOT in MS
+
+missing_MS <- compare_df %>% filter(existsMS == 0)
+
+missing_MS
+
+write.csv(missing_MS, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/missing_MS.csv')
+```
+
+## Plan: want to add pollen hosts from MS data so I can filter on pollen family. Discuss how to parse the authors out of appropriate MS column. ## DONE
+NEED TO DISCUSS: The MS pollen hosts are a mix of genus and families, randomly. Maybe we can ask MS to seperate these into two columns, or just simply make a family level column only.
 
@@
 
-## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa)
+## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa) ## WIP
 
-## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables)
+## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables). ## DONE
+DISCUSS: Turns out that we didnt cut globi bees, but we instead cut FOWLER bees. I unfiltered the fowler bees and it fixed the ratios. Now we can actually discuss data proportions
+
+### Also: Added the target pollen hosts from globi so we can compare to pollen hosts from MS
+This way we can directly compare the MS pollen hosts to the GLOBI pollen hosts. THe only issue is the above problem with a mix of taxonomic levels with the MS data
+
+---------------
+
+For next week:
+
+Check out the web scraper to fix MS names to family level
+
+Pull out data that is only in GLOBI and/or Fowler, use as group to check for old names as reason.
+
+--------------
+
+## Modeling below
+
+# (1) Model same as old, except use conclusions from MS not fowler
 
 ```{r}
+modeling_df <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% mutate(specialist = ifelse(specializationMS == 'Polylectic', 1, 0 )) %>% dplyr::select(interestBee, nBeeGLOBEE, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, FamilyDegreeMS, specialist)
 
+modeling_df
 ```
@@
 ```{r}
+calc_error_rate = function(predicted.value, true.value){
+  return(mean(true.value!=predicted.value))
+}
+records = matrix(NA, nrow=3, ncol=2)
+colnames(records) = c("train.error","test.error")
+rownames(records) = c("tree","logistic","lasso")
 
+nfold <- 10
+
+n <- nrow(modeling_df)
+
+in.trn <- sample.int(n, 0.8*n) 
+trn.cl <- modeling_df[ in.trn,]
+tst.cl <- modeling_df[-in.trn,]
+
+folds <- sample(cut(1:nrow(trn.cl), breaks=nfold, labels=FALSE))
 ```
@@
 ```{r}
+library(tree)
+library(maptree)
 
+tree <- tree(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl)
+cv <- cv.tree(tree, rand = folds, FUN = prune.misclass, K = 10)
+
+best.size.cv = min(cv$size[cv$dev == min(cv$dev)])
+
+tree.pruned <- prune.misclass(tree, best= best.size.cv)
+
+train_pred_tree <- predict(tree.pruned, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')
+test_pred_tree <- predict(tree.pruned, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')
+
+train_err <- calc_error_rate(train_pred_tree, trn.cl$specialist)
+test_err <- calc_error_rate(test_pred_tree, tst.cl$specialist)
+
+1 - train_err
+1 - test_err
 ```
+
+```{r}
+glmfit <- glm(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl, family = 'binomial')
+predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="response")
+predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="response")
+
+predicttrain_log = ifelse(predicttrain_log > 0.5, 1, 0)
+predicttest_log = ifelse(predicttest_log > 0.5, 1, 0)
+
+trainerror <- calc_error_rate(predicttrain_log, trn.cl$specialist)
+valerror <- calc_error_rate(predicttest_log, tst.cl$specialist)
+
+1 - trainerror
+1 - valerror
+```
+
+```{r}
+library(randomForest)
+
+rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="class")
+
+train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)
+
+test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+```
+# Try multi feature with best model:
+
+```{r}
+rf_fit <-  randomForest(as.factor(specialist) ~ ., data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(-specialist), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(-specialist), type="class")
+
+train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)
+
+test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+```
+
+# Begin Baysian Approaches
+
+```{r}
+modeling_df
+```
+
+```{r}
+compare_df %>% filter(interestBee == 'Agapostemon Angelicus') 
+```

### 2022-01-26 MS_data_modeling/MS_data_explore.Rmd
Further work on paper

@@
 library(plyr)
+library(caret)
 
-MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_10-26-21_ALR seltmann-colab.xlsx', sheet = 2)
+MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/UCSB Archive/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 2)
 
@@
 ```{r}
+interactions <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv')
+
+(interactions %>% filter(targetTaxonRank == 'family') %>% dplyr::select(sourceTaxonName, interactionTypeName, targetTaxonName) %>% 
+    dplyr::rename('Bee Species' = sourceTaxonName, 'Interaction' = interactionTypeName, 'Plant Family' = targetTaxonName))[c(172, 112, 300, 124 ,5),]  %>%
+  kbl() %>% kable_styling()
+
+```
+
+```{r}
 MS_data_format <- MS_data %>% mutate(interestBee = str_to_title(paste(Genus, sub(".*? ", "", Species)))) %>% 
@@
 
-MS_data_format %>% filter(!(is.na(specializationMS)), is.na(pollenHosts)) ### there
+MS_data_format <- MS_data_format %>% filter(!(is.na(specializationMS)), !is.na(pollenHosts)) ### there
+
+MS_data_format
 ```
@@
 ms_pollen_hosts <- ms_pollen_hosts[!(ms_pollen_hosts == '')]
-
-ms_pollen_hosts
 ```
@@
 ```{r}
-family_key %>% filter(is.na(pollen_host_familys))
+family_key %>% filter(is.na(pollen_host_familys) | (pollen_host_familys == 'NA'))
 ```
@@
 ```
+
+### Note, NAs are a result of there being no pollen hsots listed for a bee (should be fixed by filtering out bees with no pollen hosts), and Unknowns are a results of bees not being able to be found in catalougue of life
 
@@
 
-compare_df
+compare_df %>% filter(existsALL == 1) %>% dplyr::select(interestBee, specializationMS, FamilyDegreeMS, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
+  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 'Degree MS' = FamilyDegreeMS, 'Raw Degree GLOBI' = degreeGLOBEE, 
+                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE) %>%
+  kbl() %>% kable_styling()
 ```
@@
 
-data.frame(data_sources, counts) %>% mutate('Percent of Data' = counts / nrow(compare_df)) %>% 
+data.frame(data_sources, counts) %>% mutate('Percent of All Bee Species Documented' = counts / nrow(compare_df)) %>% 
   dplyr::rename('Data Source' = data_sources, 'Unique Species Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
@@
 odd_generalists <- compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
-  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsFamilysMS, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 == 1)
+  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5,degreeGLOBEE_T, degreeGLOBEE, pollenHostsFamilysMS, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 == 1)
 
@@
 ```{r}
+
+(odd_generalists %>% dplyr:: select(interestBee, specializationMS, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
+  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 
+                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE))[c(3,5),] %>%
+  kbl(caption = 'Odd Generalists') %>% kable_styling()
+
+```
+
+```{r}
 # Bees NOT fowler specialists but MS specialists
@@
 odd_specialists <- compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
-  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsMS, pollenHostsFamilysMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 > 3)
-
-odd_specialists
+  dplyr::select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, degreeGLOBEE_T, pollenHostsMS, pollenHostsFamilysMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 > 3)
 
 write.csv(odd_specialists, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/odd_specialists.csv')
+```
+
+```{r}
+(odd_specialists %>% dplyr:: select(interestBee, specializationMS, degreeGLOBEE_C5, degreeGLOBEE_T, pollenHostsFamilysMS, pollenHostGLOBEE) %>%
+  dplyr::rename('Bee Species' = interestBee, 'Binary Label MS' = specializationMS, 
+                'Degree Cut GLOBI' = degreeGLOBEE_C5, 'Tranformed Degree GLOBI' = degreeGLOBEE_T, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts MS' = pollenHostsFamilysMS, 'Pollen Hosts GLOBI' = pollenHostGLOBEE) )%>%
+  kbl(caption = 'Odd Specialists') %>% kable_styling()  
 ```
@@
 ```{r}
-modeling_df <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% mutate(specialist = ifelse(specializationMS == 'Polylectic', 1, 0 )) %>% dplyr::select(interestBee, nBeeGLOBEE, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, FamilyDegreeMS, specialist)
+modeling_df <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% mutate(specialist = ifelse(specializationMS == 'Polylectic', 1, 0 )) %>% dplyr::select(interestBee, nBeeGLOBEE, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, FamilyDegreeMS, specialist, pollenHostsFamilysMS, pollenHostGLOBEE)
 
 modeling_df
+ggplot(data= modeling_df, aes(x = nBeeGLOBEE, y = degreeGLOBEE_T, color = as.factor(specialist))) + geom_point() + xlim(0, 1000)
+```
+```{r}
+length(unique(word(modeling_df$interestBee, 1)))
+
+pollen_hosts_represented <- c(modeling_df$pollenHostsFamilysMS, modeling_df$pollenHostGLOBEE)
+length(unique(pollen_hosts_represented))
+```
+
+```{r}
+set.seed(123)
 ```
@@
 ```{r}
-glmfit <- glm(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl, family = 'binomial')
-predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="response")
-predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="response")
+glmfit <- cv.glmnet(modeling_df %>% dplyr::select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), as.factor(modeling_df$specialist), family = 'binomial', alpha=1, 
+          nfolds = 10)
+
+coef  <-  coef(glmfit) ## uses lowest mse lambda
+coef
+
+predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")
+predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")
 
@@
 ```
-# Try multi feature with best model:
 
 ```{r}
-rf_fit <-  randomForest(as.factor(specialist) ~ ., data = trn.cl, importance=TRUE)
+library(randomForest)
 
-predict_train <- predict(rf_fit, newdata = trn.cl %>% select(-specialist), type="class")
-predict_test <- predict(rf_fit, newdata = tst.cl %>% select(-specialist), type="class")
+rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")
 
@@
 
-# Begin Baysian Approaches
+```{r}
+library(randomForest)
+
+rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + nBeeGLOBEE, data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")
+
+train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)
+
+test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+```
+
+# Try multi feature with best model:
 
 ```{r}
+rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + degreeGLOBEE_T + degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")
+
+train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)
+
+test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+
+misclassified <- tst.cl[tst.cl$specialist == predict_test,]
+misclassified_vec <- tst.cl$specialist != predict_test
+
+paste('Percent of misclassified that were specialists:', round(sum(misclassified$specialist) / nrow(misclassified), 3))
+paste('Percent of studied that were specialists:', round(sum(modeling_df$specialist) / nrow(modeling_df), 3))
+```
+
+```{r}
+tst.cl$misclass <- as.numeric(misclassified_vec)
+splits <- c(0, 5, 10, 100, 100000)
+tst.cl$nBeeBin <- as.factor(cut(tst.cl$nBeeGLOBEE, breaks=splits))
+tst.cl %>% dplyr::select(misclass, nBeeBin) %>% group_by(nBeeBin) %>% dplyr::summarize(freq = n(), misclass_rate = mean(misclass))
+```
+
+# Try fitting to family degree:
+
+```{r}
+ggplot(data = modeling_df, aes(x= FamilyDegreeMS)) + geom_histogram()
+
+splits <- c(0, 1, 2, 5, 10, 100)
+modeling_df$MSFamilyDegreeBins <- cut(modeling_df$FamilyDegreeMS, breaks=splits)
 modeling_df
@@
 ```{r}
-compare_df %>% filter(interestBee == 'Agapostemon Angelicus') 
+calc_error_rate = function(predicted.value, true.value){
+  return(mean(true.value!=predicted.value))
+}
+records = matrix(NA, nrow=3, ncol=2)
+colnames(records) = c("train.error","test.error")
+rownames(records) = c("tree","logistic","lasso")
+
+nfold <- 10
+
+n <- nrow(modeling_df)
+
+in.trn <- sample.int(n, 0.8*n) 
+trn.cl <- modeling_df[ in.trn,]
+tst.cl <- modeling_df[-in.trn,]
+
+folds <- sample(cut(1:nrow(trn.cl), breaks=nfold, labels=FALSE))
 ```
+
+```{r}
+library(tree)
+library(maptree)
+
+tree <- tree(as.factor(MSFamilyDegreeBins) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl)
+cv <- cv.tree(tree, rand = folds, FUN = prune.misclass, K = 10)
+
+best.size.cv = min(cv$size[cv$dev == min(cv$dev)])
+
+tree.pruned <- prune.misclass(tree, best= best.size.cv)
+
+train_pred_tree <- predict(tree, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')
+test_pred_tree <- predict(tree, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type = 'class')
+
+train_err <- calc_error_rate(train_pred_tree, trn.cl$specialist)
+test_err <- calc_error_rate(test_pred_tree, tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+```
+
+```{r}
+glmfit <- cv.glmnet(modeling_df %>% dplyr::select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), as.factor(modeling_df$specialist), family = 'binomial', alpha=1, 
+          nfolds = 10)
+
+coef  <-  coef(glmfit) ## uses lowest mse lambda
+coef
+
+predicttrain_log = predict(glmfit, trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")
+predicttest_log = predict(glmfit, tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE) %>% as.matrix(), type="response")
+
+predicttrain_log = ifelse(predicttrain_log > 0.5, 1, 0)
+predicttest_log = ifelse(predicttest_log > 0.5, 1, 0)
+
+trainerror <- calc_error_rate(predicttrain_log, trn.cl$specialist)
+valerror <- calc_error_rate(predicttest_log, tst.cl$specialist)
+
+1 - trainerror
+1 - valerror
+```
+
+```{r}
+library(randomForest)
+
+rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_T + nBeeGLOBEE, data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE), type="class")
+
+train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)
+
+test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+```
+
+```{r}
+library(randomForest)
+
+rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE, nBeeGLOBEE), type="class")
+
+train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)
+
+test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+```
+
+```{r}
+library(randomForest)
+
+rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + nBeeGLOBEE, data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_C5, nBeeGLOBEE), type="class")
+
+train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)
+
+test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+```
+
+# Try multi feature with best model:
+
+```{r}
+rf_fit <-  randomForest(as.factor(specialist) ~ degreeGLOBEE_C5 + degreeGLOBEE_T + degreeGLOBEE + nBeeGLOBEE, data = trn.cl, importance=TRUE)
+
+predict_train <- predict(rf_fit, newdata = trn.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")
+predict_test <- predict(rf_fit, newdata = tst.cl %>% select(degreeGLOBEE_T, nBeeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE), type="class")
+
+train_err <- calc_error_rate(as.character(predict_train), trn.cl$specialist)
+
+test_err <- calc_error_rate(as.character(predict_test), tst.cl$specialist)
+
+1 - train_err
+1 - test_err
+```
+
+```{r}
+max(modeling_df$degreeGLOBEE_C5)
+```
+
+## Step 1: Classify bees as weird by finding misclassified bees by ML algoritem
+## Step 2: Pull family level from both data and literature
+## Step 3: Filter differences in family citations
+## Step 4: Find similarities of family discrepencies (this identifies bias in data)
+## Step 5:

### 2022-06-06 Full Bredth v Specialization Study.Rmd
genus level updates

---
title: "Full Bredth v Specialization Study"
author: "Nick R. Bachelder"
date: "4/15/2022"
output: html_document
---

## issue fixes:

# Assume non-rare bees are 'not -rare' DONE
# Remove cases where bee rarity changes across area # DONE
# Change common catagory to 'not-rare' (assumption: fowler is exhustive) # DONE
# location can be tossed due to data limitation, use broad scale east, west, central together.
# Do it at the genus level for plants

### aceae for families if we clean out family names for plants
### use tp for source

### Question: For and common rare bees in the easter, western, and central united states respectively, is the specialization indicative of limited diet bredth?
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
compare_df <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/compare_df.csv') %>% dplyr::select(-X)

compare_df %>% filter(degreeMS > degreeMSGenus)
```

### lets experiment with all bees at FAMILY LEVEL

```{r}
monolectic <- compare_df %>% filter(specializationMS %in% c('Monolectic', 'Oligolectic'))
polylectic <- compare_df %>% filter(specializationMS == 'Polylectic')

monolectic
polylectic
```

### Note: correlation for polylectic

```{r}
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE, method=c("pearson"))
```

```{r}
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE_C5, method=c("pearson"))
```

## do at genus level

```{r}

```

```{r}

```

```
```

### lets experiment with all bees at GENUS LEVEL

```{r}
compare_df
```

### Note: monolectic all have degreeMS = 1, so cant measure correlation. polylectic all similar

```{r}
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE, method=c("pearson"))
```

```{r}
cor.test(monolectic$degreeMS, monolectic$degreeGLOBEE_C5, method=c("pearson"))
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE_C5, method=c("pearson"))
```

### Note: No correlation between MSdegree and GLOBEE_T degree for  monolectic, but one for polylectic

```{r}
cor.test(monolectic$degreeMS, monolectic$degreeGLOBEE_T, method=c("pearson"))
cor.test(polylectic$degreeMS, polylectic$degreeGLOBEE_T, method=c("pearson"))
```

```{r}
monolectic %>% ggplot(aes(x = degreeMS, y = degreeGLOBEE_T)) + geom_point()
```

```{r}
polylectic %>% ggplot(aes(x = degreeMS, y = degreeGLOBEE_T)) + geom_point()
```

### 2022-06-06 MS_data_modeling/MS_data_explore.Rmd
genus level updates

--- 2021-10-15 MS_data_modeling/MS_data_explore.Rmd
@@
 library(kableExtra)
+library(rcol)
+library(plyr)
+library(caret)
 
-MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_09-26-21_ALR_seltmann-colab.xlsx', sheet = 2)
+MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/UCSB Archive/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 2)
 
-globee <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
-  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+globee <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteRaw') %>%
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
-globee_5cut <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteC5') %>%
-  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+globee_5cut <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteC5') %>%
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
-globee_T <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteW') %>%
-  select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+globee_T <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteW') %>%
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
@@
    mutate(interestBee = str_to_title(interestBee)) %>%
-  select(interestBee, west_specialist, east_specialist, cent_specialist, specialist)
+  dplyr::select(interestBee, west_specialist, east_specialist, cent_specialist, specialist, Status_E, Status_W, Status_C) %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+globeeGenus <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteRawGenus') %>%
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+
+globee_5cutGenus <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteC5Genus') %>%
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+
+globee_TGenus <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteWGenus') %>%
+  dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
+```
+
+```{r}
+interactions <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv')
+
+(interactions %>% filter(targetTaxonRank == 'family') %>% dplyr::select(sourceTaxonName, interactionTypeName, targetTaxonName) %>% 
+    dplyr::rename('Bee Species' = sourceTaxonName, 'Interaction' = interactionTypeName, 'Plant Family' = targetTaxonName))[c(172, 112, 300, 124 ,5),]  %>%
+  kbl() %>% kable_styling()
+
 ```
@@
 MS_data_format <- MS_data %>% mutate(interestBee = str_to_title(paste(Genus, sub(".*? ", "", Species)))) %>% 
-  rename(degreeMS = 'Pollen host family #', specializationMS = 'Redefined pollen diet breadth', pollenHosts = 'Pollen hosts') %>%
-  select(interestBee, pollenHosts, degreeMS, specializationMS, References) %>%
-  filter(!is.na(degreeMS)) %>% arrange(interestBee) %>% mutate(pollenHosts = str_to_title(gsub("[A-Za-z]*[^A-Za-z/, ]\\S*\\s*", "", pollenHosts)))
+  dplyr::rename(degreeMS = 'Pollen host family #', degreeMSGenus = 'Pollen host genus #', specializationMS = 'Redefined pollen diet breadth', pollenHosts = 'Pollen hosts') %>%
+  dplyr::select(interestBee, pollenHosts, degreeMS, degreeMSGenus, specializationMS, References) %>%
+  filter(!is.na(specializationMS)) %>% arrange(interestBee) %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+MS_data_format <- MS_data_format %>% filter(!(is.na(specializationMS)), !is.na(pollenHosts)) ### there
 
@@
 
+### create family level column by pulling
+
 ```{r}
-pollenHostGLOBEE <- aggregate(interact~interestBee, globee_5cut, FUN=toString) %>% mutate(interestBee = str_to_title(interestBee)) %>% 
-  rename('pollenHostGLOBEE' = interact)
+cp_nu_search(q='Helianthus', status = "accepted")$result$classification
+```
+
+```{r}
+MS_data_format$pollenHosts <-  gsub("\\?.*", "", MS_data_format$pollenHosts)
+MS_data_format$pollenHosts <-  gsub("\\..*", "", MS_data_format$pollenHosts)
+
+ms_pollen_hosts = na.omit(unique(
+
+  sub("\\/.*", "", word(unlist(str_split(unique(MS_data_format$pollenHosts), pattern = ', ')), 1))
+
+  ))
+
+## pulls all unique maximum level taxonomic names from MS pollen hosts
+ms_pollen_hosts <- ms_pollen_hosts[!(ms_pollen_hosts == '')]
+```
+
+```{r}
+library(rcol)
+library(pbapply)
+
+find_family <- function(name) {
+  try <- try(cp_nu_search(q=name, status = "accepted"), silent=TRUE)
+  if ("try-error" %in% class(try)) {
+    label <- 'Unknown'
+  }
+  try <- try(str_to_title(tolower(
+    (rbind( data.frame(try$result$classification[1]),
+        data.frame(try$result$classification[2]),
+       data.frame(try$result$classification[3])) %>% filter(rank == 'family'))[1,]$label
+    )), silent=TRUE)
+  if ("try-error" %in% class(try)) {
+    label <- 'Unknown'
+  }
+  else {
+    label <- try
+  }
+  return(label)
+}
+
+familys <- pblapply(ms_pollen_hosts, find_family)
+```
+
+```{r}
+familys <- unlist(familys)
+family_key <- data.frame(cbind(ms_pollen_hosts, familys)) %>% `colnames<-`(c('max_taxons', 'pollen_host_familys'))
+
+family_key
+```
+
+## Make the below two entries have 0 rows with Avery's help
+```{r}
+family_key %>% filter(is.na(pollen_host_familys) | (pollen_host_familys == 'NA'))
+```
+
+```{r}
+family_key %>% filter(pollen_host_familys == 'Unknown', grepl( ',', max_taxons, fixed = TRUE))
+```
+
+## make sure the below has no family name synonyms. Take out the unknownand NA values until we fix them with avery.
+
+```{r}
+pollen_family_MS <- MS_data_format %>% select(interestBee, pollenHosts) %>% separate_rows(pollenHosts, sep=', ') %>% mutate(max_taxons = word(pollenHosts, 1)  ) %>% left_join(family_key, by = 'max_taxons') %>% select(interestBee, pollen_host_familys) %>%
+  group_by(interestBee) %>% 
+  dplyr::summarise(pollen_host_familys = paste(unique(as.vector(pollen_host_familys)), collapse = ', '))
+
+MS_data_format <- MS_data_format %>% left_join(pollen_family_MS, by = 'interestBee') 
+```
+
+### Note, NAs are a result of there being no pollen hsots listed for a bee (should be fixed by filtering out bees with no pollen hosts), and Unknowns are a results of bees not being able to be found in catalougue of life
+
+```{r}
+avery_send <- MS_data_format %>% select(interestBee, pollenHosts, pollen_host_familys) %>% mutate(ifelse(pollen_host_familys == 'NA', NA, pollen_host_familys))
+
+write.csv(avery_send, "/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Family_Level_Transform.csv")
+
+avery_send
+```
+
+### Clean GLOBEE data
+
+```{r}
+pollenHostGLOBEE <- aggregate(interactFamily~interestBee, globee_5cut, FUN=toString) %>% mutate(interestBee = str_to_title(interestBee)) %>% 
+  dplyr::rename('pollenHostGLOBEEFamily' = interactFamily) %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
 ## this is plant interactions for each bee genus according to globi
 
-globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE = degree) %>% 
-  select(interestBee, degreeGLOBEE) %>% unique() %>%
-  arrange(interestBee) %>% merge(pollenHostGLOBEE, by = 'interestBee')
+pollenHostGLOBEEGenus <- aggregate(interactGenus~interestBee, globee_5cutGenus, FUN=toString) %>% mutate(interestBee = str_to_title(interestBee)) %>% 
+  dplyr::rename('pollenHostGLOBEEGenus' = interactGenus) %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+## this is plant interactions for each bee genus according to globi
 
-globee_5cut_format <- globee_5cut %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE_C5 = degree) %>% 
-  select(interestBee, degreeGLOBEE_C5) %>% unique() %>%
-  arrange(interestBee)
+globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEEFamily = degreeFamily) %>% 
+  dplyr::select(interestBee, degreeGLOBEEFamily, number_of_citation_for_bee_species) %>% unique() %>%
+  arrange(interestBee) %>% merge(pollenHostGLOBEE, by = 'interestBee') %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
 
-globee_T_format <- globee_T %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE_T = degree) %>% 
-  select(interestBee, degreeGLOBEE_T) %>% unique() %>%
-  arrange(interestBee)
+globee_5cut_format <- globee_5cut %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_C5Family = degreeC5Family) %>% 
+  dplyr::select(interestBee, degreeGLOBEE_C5Family) %>% unique() %>%
+  arrange(interestBee)%>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+globee_T_format <- globee_T %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_TFamily = degreeWFamily) %>% 
+  dplyr::select(interestBee, degreeGLOBEE_TFamily) %>% unique() %>%
+  arrange(interestBee)%>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+globee_formatGenus <- globeeGenus %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEEGenus = degreeGenus) %>% 
+  dplyr::select(interestBee, degreeGLOBEEGenus, number_of_citation_for_bee_species) %>% unique() %>%
+  arrange(interestBee) %>% merge(pollenHostGLOBEEGenus, by = 'interestBee') %>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+globee_5cut_formatGenus <- globee_5cutGenus %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_C5Genus = degreeC5Genus) %>% 
+  dplyr::select(interestBee, degreeGLOBEE_C5Genus) %>% unique() %>%
+  arrange(interestBee)%>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+globee_T_formatGenus <- globee_TGenus %>% mutate(interestBee = str_to_title(interestBee)) %>% dplyr::rename(degreeGLOBEE_TGenus = degreeWGenus) %>% 
+  dplyr::select(interestBee, degreeGLOBEE_TGenus) %>% unique() %>%
+  arrange(interestBee)%>% 
+  filter(grepl('^\\w+\\s\\w+$', interestBee))
 ```
@@
 compare_df <- merge(compare_df, globee_T_format, by = c('interestBee'), all = T) 
+compare_df <- merge(compare_df, globee_formatGenus, by = c('interestBee', 'number_of_citation_for_bee_species'), all = T)  
+compare_df <- merge(compare_df, globee_5cut_formatGenus, by = c('interestBee'), all = T) 
+compare_df <- merge(compare_df, globee_T_formatGenus, by = c('interestBee'), all = T)
 
-compare_df <- compare_df %>% merge(fowler %>% rename(specialistFOWLER = specialist) %>% 
-                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% unique() %>%
-  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(degreeMS), 0, 1), 
+compare_df <- compare_df %>% merge(fowler %>% dplyr::rename(specialistFOWLER = specialist) %>% 
+                                     dplyr::select(interestBee, specialistFOWLER, Status_E, Status_W, Status_C), by = c('interestBee'), all = T) %>% unique() %>%
+  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEEFamily), 0, 1), existsMS = ifelse(is.na(specializationMS), 0, 1), 
          existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
-         existsALL = existsGLOBEE*existsMS,
-         pollenHostGLOBEE = str_to_title(pollenHostGLOBEE)) %>%
+         existsALL = existsGLOBEE*existsMS) %>%
   mutate(specialistFOWLER = ifelse(is.na(specialistFOWLER), 0 , specialistFOWLER))%>%
-  rename(pollenHostsMS = pollenHosts,ReferencesMS = References) %>%
-  select(interestBee, specializationMS, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, specialistFOWLER, 
-         existsGLOBEE, existsMS, existsFOWLER, existsALL, pollenHostsMS, pollenHostGLOBEE, ReferencesMS, degreeMS)
+  dplyr::rename(pollenHostsMS = pollenHosts , pollenHostsFamilysMS = pollen_host_familys,ReferencesMS = References) %>%
+  rowwise() %>%
+  dplyr::mutate(pollenHostsFamilysMS = ifelse(pollenHostsFamilysMS == 'NA', NA, pollenHostsFamilysMS) ) %>%
+  dplyr::mutate(FamilyDegreeMS = ifelse(!is.na(pollenHostsFamilysMS), length(strsplit(pollenHostsFamilysMS, split = ', ')[[1]]), 0  ) )
 
-compare_df %>% filter(existsALL == 0)
+final <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% 
+  dplyr::rename(interest_bee_species = interestBee, number_of_citations_for_bee_species_GloBi = number_of_citation_for_bee_species, degree_raw_GloBi_family = degreeGLOBEEFamily,
+                pollen_hosts_GloBi_family = pollenHostGLOBEEFamily, pollen_host_genus_MS = pollenHostsMS, degree_MS_family = degreeMS, degree_MS_Genus = degreeMSGenus,
+                specialization_MS = specializationMS, References_MS = ReferencesMS, pollen_host_family_MS = pollenHostsFamilysMS, 
+                degree_cut_5_GloBi_family = degreeGLOBEE_C5Family, degree_transformed_GloBi_family = degreeGLOBEE_TFamily, 
+                degree_raw_GloBi_genus = degreeGLOBEEGenus, pollen_hosts_GloBi_genus = pollenHostGLOBEEGenus, degree_cut_5_GloBi_genus = degreeGLOBEE_C5Genus,
+                degree_transformed_GloBi_genus = degreeGLOBEE_TGenus, specialist_FOWLER = specialistFOWLER) %>% 
+  dplyr::select(interest_bee_species, number_of_citations_for_bee_species_GloBi, degree_raw_GloBi_family, degree_cut_5_GloBi_family, degree_transformed_GloBi_family,
+                pollen_hosts_GloBi_family, degree_MS_family, pollen_host_family_MS, degree_raw_GloBi_genus, degree_cut_5_GloBi_genus, degree_transformed_GloBi_genus,
+                pollen_hosts_GloBi_genus, pollen_host_genus_MS, specialization_MS) %>% 
+           mutate(pollen_host_family_MS = str_remove(pollen_hosts_GloBi_family, "NA, "),
+                  pollen_hosts_GloBi_family = str_to_title(pollen_hosts_GloBi_family), pollen_host_family_MS =str_to_title(pollen_host_family_MS),
+                  pollen_hosts_GloBi_genus = str_to_title(pollen_hosts_GloBi_genus))
+
+final
+
+write.csv(final, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/specialization_degree_df')
 ```
 
-### Add individual citation cuts in addition to raw GLOBI
-### Figure out what we want to define as 'specialist' MS data
-
-```{r}
-only_globi <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 0, existsFOWLER == 0) %>% nrow()
-globi_fowler <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 0, existsFOWLER == 1) %>% nrow()
-globi_MS <- compare_df %>% filter(existsGLOBEE == 1, existsMS == 1, existsFOWLER == 0) %>% nrow()
-MS_fowler <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 1, existsFOWLER == 1) %>% nrow()
-fowler_only <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 0, existsFOWLER == 1) %>% nrow()
-MS_only <- compare_df %>% filter(existsGLOBEE == 0, existsMS == 1, existsFOWLER == 0) %>% nrow()
-all_sources <- compare_df %>% filter(existsALL == 1) %>% nrow()
-
-counts <- c(only_globi,fowler_only,MS_only, globi_MS, globi_fowler, MS_fowler, all_sources, nrow(compare_df))
-data_sources <- c('Only in GLOBI', 'Only in Fowler', 'Only in MS', 'GLOBI and MS', 'GLOBI and Fowler', 'MS and Fowler', 'All Sources', 'Total')
-
-data.frame(data_sources, counts) %>% mutate('Percent of Data' = counts / nrow(compare_df)) %>% 
-  rename('Data Source' = data_sources, 'Unique Genus Count' = counts) %>% kbl(caption = "Bee Data Sources Summary") %>%
-  kable_styling()
-```
-
-```{r}
-compare_df %>% filter(existsALL == 1) %>% group_by(specializationMS) %>% summarize('Mean GLOBI Degree'  = mean(degreeGLOBEE),
-                                                                                   'GLOBI Degree IQR' = paste('[', round(quantile(degreeGLOBEE, 0.25),
-                                                                                                                         2), 
-                                                                                                                  ',', round(quantile(degreeGLOBEE, 0.75), 2),
-                                                                                                            ']'),
-                                                                                   'Mean GLOBI C5 Degree' = mean(degreeGLOBEE_C5),
-                                                                                   'GLOBI C5 Degree IQR' = paste('[', round(quantile(degreeGLOBEE_C5, 0.25), 2),
-                                                                                                                  ',', round(quantile(degreeGLOBEE_C5, 0.75), 2),
-                                                                                                            ']'),
-                                                                                   'Mean GLOBI T Degree' = mean(degreeGLOBEE_T),
-                                                                                   'GLOBI T Degree IQR' = paste('[', round(quantile(degreeGLOBEE_T, 0.25), 2),
-                                                                                                                  ',', round(quantile(degreeGLOBEE_T, 0.75), 2),
-                                                                                                            ']'),
-                                                                                   'Percent Fowler Specialist' = mean(specialistFOWLER)
-) %>% kbl(caption = 'Plant Family Degrees by MS Specialization') %>%
-  kable_styling()
-```
-
-```{r}
-# Bees fowler specialists but not MS specialists
-
-compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE < 3)
-```
-
-```{r}
-# Bees NOT fowler specialists but MS specialists
-
-compare_df %>% filter(specialistFOWLER == 0, specializationMS  != 'Polylectic') %>% 
-  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHostsMS, pollenHostGLOBEE, ReferencesMS) %>% filter(degreeGLOBEE_C5 > 5)
-```
-
-## Plan: want to add pollen hosts so I can filter on pollen family. Discuss how to parse the authors out of appropriate MS column. ## DONE
-
-## Cleaning up source and pollen columns (removin author abbreviation)   ### DONE
-
-## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa) ## WIP
-
-## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables). ## DONE: Turns out that we didnt cut globi bees, but we instead cut FOWLER bees.
-
-### Also: Added the target pollen hosts from globi so we can compare to pollen hosts from MS
-
-```{r}
-
-```
-
-```{r}
-
-```
-
-```{r}
-
-```
+---------------

### 2022-06-06 modeling_data/GLOBI_bee_update_clean.Rmd
folder organization
genus level updates

--- 2021-09-30 GLOBI_bee_update_clean.Rmd
@@
 
-plant_source_df <- plant_source_df %>% rename(sourceCitation = targetCitation, sourceNamespace = targetNamespace, 
+plant_source_df <- plant_source_df %>% dplyr::rename(sourceCitation = targetCitation, sourceNamespace = targetNamespace, 
                                               sourceArchiveURI = targetArchiveURI, sourceDOI = targetDOI,
@@
 
-## Result
+# Export to csv file
 ```{r}
-interactions
+write.csv(interactions_dup, "interactions_dup.csv")
 ```
-
-# Export to csv.gz file
-```{r}
-write.csv(interactions, "generated_data/interactions.csv")
-write.csv(interactions_dup, "generated_data/interactions_dup.csv")
-```

### 2022-07-20 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep.Rmd
genus level push

@@
 library(data.table)
+
 ```
@@
 options(max.print=1000000)
-interactions_b <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv') %>% select(-X)
+interactions_b <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv') %>% dplyr::select(-X)
 
@@
 
+```{r}
+interactions_b[1:5,]
+```
+
 Define beeCites with varying criteria
@@
 ```{r, fig.width=5,fig.height=2}
-beeCiteNoFilter <- interactions_b %>% select(sourceTaxonSpeciesName, targetTaxonFamilyName) %>% dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% group_by(interestBee, interact) %>% count() %>% ungroup() %>% filter(interestBee != "", interact != "")
+beeCiteNoFilter <- interactions_b %>% dplyr::select(sourceTaxonSpeciesName,targetTaxonFamilyName) %>% 
+  dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% dplyr::group_by(interestBee, interact) %>% dplyr::count() %>% ungroup() %>% 
+  filter(interestBee != "", interact != "")
 
-TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(n)) %>% ungroup()
+TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% dplyr::summarise(n_Bee = sum(n)) %>% ungroup()
 TotalCitations <- sum(TotalCitations$n_Bee)
 
-beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
-beeCiteNoFilter <- beeCiteNoFilter %>% summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()
+beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% dplyr::summarise(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
+beeCiteNoFilter <- beeCiteNoFilter %>% dplyr::summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup() %>% unique()
 
-beeIntDegree <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
-beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(n))
+beeIntDegree <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count() 
+beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique()
 
-beeIntDegreeC10 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
+beeIntDegreeC10 <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
 beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
 beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
-beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 10 citations to contribute to degree
+beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 10 citations to contribute to degree
 
-beeIntDegreeC5 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
+beeIntDegreeC5 <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
 beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
 beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
-beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 5 citations to contribute to degree
+beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 5 citations to contribute to degree
 
-beeIntDegreeW <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
+beeIntDegreeW <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
 beeIntDegreeW$numcitations <- beeCiteNoFilter$n
@@
 beeIntDegreeW$n <- (beeIntDegreeW$numcitations / beeIntDegreeW$n_Bee)^2
-beeIntDegreeW <- beeIntDegreeW %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## weighted degree
+beeIntDegreeW <- beeIntDegreeW %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## weighted degree
 
-beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type = 'inner')
+beeCite <-  left_join(beeCiteNoFilter, beeIntDegree, by = 'interestBee')
 
-beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee",  type = 'inner')
+beeCiteC10 <-  left_join(beeCiteNoFilter, beeIntDegreeC10, by = 'interestBee')
 
-beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee",  type = 'inner')
+beeCiteC5 <-  left_join(beeCiteNoFilter, beeIntDegreeC5, by = 'interestBee')
 
-beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee",  type = 'inner')
+beeCiteW <-  left_join(beeCiteNoFilter, beeIntDegreeW, by = 'interestBee')
 
-beeCiteW
+beeCiteNoFilterGenus <- interactions_b %>% dplyr::select(sourceTaxonSpeciesName, targetTaxonGenusName) %>% 
+  dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonGenusName) %>% group_by(interestBee, interact) %>% dplyr::count() %>% ungroup() %>% 
+  filter(interestBee != "", interact != "")
+
+TotalCitationsGenus <- beeCiteNoFilterGenus %>% group_by(interestBee) %>% dplyr::summarise(n_Bee = sum(n)) %>% ungroup()
+TotalCitationsGenus <- sum(TotalCitationsGenus$n_Bee)
+
+beeCiteNoFilterGenus <- beeCiteNoFilterGenus %>% group_by(interestBee) %>% dplyr::summarise(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
+beeCiteNoFilterGenus <- beeCiteNoFilterGenus %>% dplyr::summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()
+
+beeIntDegreeGenus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
+beeIntDegreeGenus <- beeIntDegreeGenus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique()
+
+beeIntDegreeC10Genus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
+beeIntDegreeC10Genus$numcitations <- beeCiteNoFilterGenus$n
+beeIntDegreeC10Genus <- beeIntDegreeC10Genus %>% filter(numcitations >= 10)
+beeIntDegreeC10Genus <- beeIntDegreeC10Genus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 10 citations to contribute to degree
+
+beeIntDegreeC5Genus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
+beeIntDegreeC5Genus$numcitations <- beeCiteNoFilterGenus$n
+beeIntDegreeC5Genus <- beeIntDegreeC5Genus %>% filter(numcitations >= 5)
+beeIntDegreeC5Genus <- beeIntDegreeC5Genus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 5 citations to contribute to degree
+
+beeIntDegreeWGenus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
+beeIntDegreeWGenus$numcitations <- beeCiteNoFilterGenus$n
+beeIntDegreeWGenus$n_Bee <- beeCiteNoFilterGenus$n_Bee
+beeIntDegreeWGenus$n <- (beeIntDegreeWGenus$numcitations / beeIntDegreeWGenus$n_Bee)^2
+beeIntDegreeWGenus <- beeIntDegreeWGenus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## weighted degree
+
+beeCiteGenus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeGenus, by = 'interestBee')
+
+beeCiteC10Genus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeC10Genus, by = 'interestBee')
+
+beeCiteC5Genus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeC5Genus, by = 'interestBee')
+
+beeCiteWGenus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeWGenus, by = 'interestBee')
+
+beeCite[1:10,]
+```
+
+```{r}
+beeCiteGenus <- beeCiteGenus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
+                        bee_proportion_of_all_bee_citations = BeeFreq, degreeGenus = degree)
+
+beeCite <- beeCite %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
+                        bee_proportion_of_all_bee_citations = BeeFreq, degreeFamily = degree)
+
+beeCiteC5Genus <- beeCiteC5Genus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
+                        bee_proportion_of_all_bee_citations = BeeFreq, degreeC5Genus = degree)
+
+beeCiteC5 <- beeCiteC5 %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
+                        bee_proportion_of_all_bee_citations = BeeFreq, degreeC5Family = degree)
+
+beeCiteWGenus <- beeCiteWGenus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
+                        bee_proportion_of_all_bee_citations = BeeFreq, degreeWGenus = degree)
+beeCiteW <- beeCiteW %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
+                        bee_proportion_of_all_bee_citations = BeeFreq, degreeWFamily = degree)
+
+beeCiteW[1:5,]
 ```
@@
 western_table$west_specialist <- 1
-western_test <- western_table %>% dplyr::select(interestBee, west_specialist) %>% unique()
+western_test <- western_table %>% dplyr::select(interestBee, west_specialist, Status) %>% dplyr::rename('Status_W' = 'Status') %>% unique()
 
 eastern_table$east_specialist <- 1
-eastern_test <- eastern_table %>% dplyr::select(interestBee, east_specialist) %>% unique()
+eastern_test <- eastern_table %>% dplyr::select(interestBee, east_specialist, Status)%>% dplyr::rename('Status_E'= 'Status') %>% unique()
 
 central_table$cent_specialist <- 1
-central_test <- central_table %>% dplyr::select(interestBee, cent_specialist) %>% unique()
+central_test <- central_table %>% dplyr::select(interestBee, cent_specialist, Status)%>% dplyr::rename('Status_C'= 'Status') %>% unique()
 
@@
 table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)
-table_test[is.na(table_test)] <- 0
 
@@
 
-fit = glm(specialist ~ n_Bee + degree, data = plotdf, family = 'binomial')
-
-ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
-  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 40))+ geom_jitter()+ scale_color_discrete(name = "Specialist") + geom_smooth()
-
 write.csv(plotdf, 'BeeCiteRaw')
@@
 plotdf$specialist[is.na(plotdf$specialist)] <- 0
-
-ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist)), text = paste("Bee:", interestBee)) +
-  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Unique Plants Interacted With (Degree)')) + ggtitle(paste('Bee Interaction Diversity (Minimum 5 Citations) and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 15))+ geom_jitter() + scale_color_discrete(name = "Specialist")+ geom_smooth(se = FALSE)
 
@@
 
-ggplot(plotdf, aes(x= n_Bee, y= degree, color = as.factor(specialist))) +
-  geom_point() + xlab(paste('Number of Citations of Bee Genus')) + ylab(paste('Function of Degree')) + ggtitle(paste('Function of Degree and Bee Citation Count')) + scale_x_continuous(limits = c(0, 200)) + scale_y_continuous(limits = c(0, 1)) + geom_smooth(se = FALSE)+ scale_color_discrete(name = "Specialist")
-
 write.csv(plotdf, 'BeeCiteW')
 ```
+
+```{r}
+plotdf <- merge(beeCiteGenus, table_test, by = 'interestBee', all.x = TRUE)
+plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
+plotdf$specialist <- as.integer(plotdf$specialist > 0)
+plotdf$specialist[is.na(plotdf$specialist)] <- 0
+
+write.csv(plotdf, 'BeeCiteRawGenus')
+```
+
+```{r}
+plotdf <- merge(beeCiteC5Genus, table_test, by = 'interestBee', all.x = TRUE)
+plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
+plotdf$specialist <- as.integer(plotdf$specialist > 0)
+plotdf$specialist[is.na(plotdf$specialist)] <- 0
+
+write.csv(plotdf, 'BeeCiteC5Genus')
+```
+
+```{r}
+plotdf <- merge(beeCiteWGenus, table_test, by = 'interestBee', all.x = TRUE)
+plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
+plotdf$specialist <- as.integer(plotdf$specialist > 0)
+plotdf$specialist[is.na(plotdf$specialist)] <- 0
+
+write.csv(plotdf, 'BeeCiteWGenus')
+```

### 2022-07-20 MS_data_modeling/MS_data_explore.Rmd
genus level push

@@
   filter(grepl('^\\w+\\s\\w+$', interestBee))
+
+globee_formatGenus %>% filter(interestBee == 'Andrena Accepta')
+globee_5cut_formatGenus %>% filter(interestBee == 'Andrena Accepta')
 ```
@@
 ```{r}
-compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), all = T)  
-compare_df <- merge(compare_df, globee_5cut_format, by = c('interestBee'), all = T) 
-compare_df <- merge(compare_df, globee_T_format, by = c('interestBee'), all = T) 
-compare_df <- merge(compare_df, globee_formatGenus, by = c('interestBee', 'number_of_citation_for_bee_species'), all = T)  
-compare_df <- merge(compare_df, globee_5cut_formatGenus, by = c('interestBee'), all = T) 
-compare_df <- merge(compare_df, globee_T_formatGenus, by = c('interestBee'), all = T)
+compare_df <- merge(globee_format, MS_data_format, by = c('interestBee'), type = 'all')  
+compare_df <- merge(compare_df, globee_5cut_format, by = c('interestBee'), type = 'all') 
+compare_df <- merge(compare_df, globee_T_format, by = c('interestBee'), type = 'all') 
+compare_df <- merge(compare_df, globee_formatGenus, by = c('interestBee', 'number_of_citation_for_bee_species'), type = 'all') 
+compare_df <- merge(compare_df, globee_5cut_formatGenus, by = c('interestBee'), type = 'all') 
+compare_df <- merge(compare_df, globee_T_formatGenus, by = c('interestBee'), type = 'all')
 
@@
 
+compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% dplyr::select(interestBee, degreeGLOBEEGenus, degreeGLOBEE_C5Genus) %>%
+  filter(!is.na(degreeGLOBEE_C5Genus), is.na(degreeGLOBEEGenus))
+
 final <- compare_df %>% filter(existsMS == 1, existsGLOBEE == 1) %>% 
   dplyr::rename(interest_bee_species = interestBee, number_of_citations_for_bee_species_GloBi = number_of_citation_for_bee_species, degree_raw_GloBi_family = degreeGLOBEEFamily,
-                pollen_hosts_GloBi_family = pollenHostGLOBEEFamily, pollen_host_genus_MS = pollenHostsMS, degree_MS_family = degreeMS, degree_MS_Genus = degreeMSGenus,
+                visitation_hosts_GloBi_family = pollenHostGLOBEEFamily, pollen_host_genus_MS = pollenHostsMS, degree_MS_family = degreeMS, degree_MS_Genus = degreeMSGenus,
                 specialization_MS = specializationMS, References_MS = ReferencesMS, pollen_host_family_MS = pollenHostsFamilysMS, 
                 degree_cut_5_GloBi_family = degreeGLOBEE_C5Family, degree_transformed_GloBi_family = degreeGLOBEE_TFamily, 
-                degree_raw_GloBi_genus = degreeGLOBEEGenus, pollen_hosts_GloBi_genus = pollenHostGLOBEEGenus, degree_cut_5_GloBi_genus = degreeGLOBEE_C5Genus,
-                degree_transformed_GloBi_genus = degreeGLOBEE_TGenus, specialist_FOWLER = specialistFOWLER) %>% 
+                degree_raw_GloBi_genus = degreeGLOBEEGenus, visitation_hosts_GloBi_genus = pollenHostGLOBEEGenus, degree_cut_5_GloBi_genus = degreeGLOBEE_C5Genus,
+                degree_transformed_GloBi_genus = degreeGLOBEE_TGenus, rarity_east_fowler = Status_E, rarity_central_fowler = Status_C, rarity_west_fowler = Status_W,
+                degree_MS_genus = degreeMSGenus) %>% 
   dplyr::select(interest_bee_species, number_of_citations_for_bee_species_GloBi, degree_raw_GloBi_family, degree_cut_5_GloBi_family, degree_transformed_GloBi_family,
-                pollen_hosts_GloBi_family, degree_MS_family, pollen_host_family_MS, degree_raw_GloBi_genus, degree_cut_5_GloBi_genus, degree_transformed_GloBi_genus,
-                pollen_hosts_GloBi_genus, pollen_host_genus_MS, specialization_MS) %>% 
-           mutate(pollen_host_family_MS = str_remove(pollen_hosts_GloBi_family, "NA, "),
-                  pollen_hosts_GloBi_family = str_to_title(pollen_hosts_GloBi_family), pollen_host_family_MS =str_to_title(pollen_host_family_MS),
-                  pollen_hosts_GloBi_genus = str_to_title(pollen_hosts_GloBi_genus))
+                visitation_hosts_GloBi_family, degree_MS_family, pollen_host_family_MS, degree_raw_GloBi_genus, degree_cut_5_GloBi_genus, 
+                degree_transformed_GloBi_genus, degree_MS_genus,
+                visitation_hosts_GloBi_genus, pollen_host_genus_MS, specialization_MS, rarity_east_fowler, rarity_central_fowler, rarity_west_fowler) %>% 
+           mutate(pollen_host_family_MS = str_remove(visitation_hosts_GloBi_family, "NA, "),
+                  visitation_hosts_GloBi_family = str_to_title(visitation_hosts_GloBi_family), pollen_host_family_MS =str_to_title(pollen_host_family_MS),
+                  visitation_hosts_GloBi_genus = str_to_title(visitation_hosts_GloBi_genus))
 
@@
 
-write.csv(final, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/specialization_degree_df')
+write.csv(final, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/visitation_specialization_df')
 ```

### 2022-07-21 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep.Rmd
changing path names

@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_knit$set(root.dir = "/Users/colleen/Dropbox/Fall_2014/ucsb/Bee-Specialization-Modeling/")
+knitr::opts_knit$set(error=T)
 
@@
 library(plotly)
-library(dplyr)
+library(tidyverse)
 library(data.table)
@@
 options(max.print=1000000)
-interactions_b <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv') %>% dplyr::select(-X)
+interactions_b <- read.csv('modeling_data/interactions_dup.csv')
 
@@
 ```{r, fig.width=5,fig.height=2}
-beeCiteNoFilter <- interactions_b %>% dplyr::select(sourceTaxonSpeciesName,targetTaxonFamilyName) %>% 
+beeCiteNoFilter <- interactions_b %>% 
+  dplyr::select(sourceTaxonSpeciesName,targetTaxonFamilyName) %>% 
   dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% dplyr::group_by(interestBee, interact) %>% dplyr::count() %>% ungroup() %>% 
@@
 
-write.csv(table_test, 'fowler')
+write_csv(table_test, 'modeling_data/fowler')
 ```
@@
 
-write.csv(plotdf, 'BeeCiteRaw')
+write_csv(plotdf, 'modeling_data/BeeCiteRaw')
 ```
@@
 
-write.csv(plotdf, 'BeeCiteC5')
+write_csv(plotdf, 'modeling_data/BeeCiteC5')
 ```
@@
 
-write.csv(plotdf, 'BeeCiteW')
+write_csv(plotdf, 'modeling_data/BeeCiteW')
 ```
@@
 
-write.csv(plotdf, 'BeeCiteRawGenus')
+write_csv(plotdf, 'modeling_data/BeeCiteRawGenus')
 ```
@@
 
-write.csv(plotdf, 'BeeCiteC5Genus')
+write_csv(plotdf, 'modeling_data/BeeCiteC5Genus')
 ```
@@
 
-write.csv(plotdf, 'BeeCiteWGenus')
+write_csv(plotdf, 'modeling_data/BeeCiteWGenus')
 ```

### 2022-07-21 MS_data_modeling/MS_data_explore.Rmd
changing path names

@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_knit$set(root.dir = "/Users/colleen/Dropbox/Fall_2014/ucsb/Bee-Specialization-Modeling/")
+knitr::opts_knit$set(error=T)
+
 library(lifecycle)
 library(rvest)
-library(dplyr)
-library(tidyr)
+library(tidyverse)
 library(plotly)
-library(dplyr)
 library(data.table)
@@
 library(caret)
+```
 
-MS_data <- readxl::read_xlsx('/Users/nickbachelder/Desktop/UCSB Archive/Fall 2021/Data for Bees/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 2)
+```{r}
+getwd()
 
-globee <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteRaw') %>%
+MS_data <- readxl::read_xlsx('modeling_data/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 2)
+
+globee <- read_csv('modeling_data/BeeCiteRaw') %>%
   dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
-globee_5cut <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteC5') %>%
+globee_5cut <- read_csv('modeling_data/BeeCiteC5') %>%
   dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
-globee_T <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteW') %>%
+globee_T <- read_csv('modeling_data/BeeCiteW') %>%
   dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
-fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/fowler') %>%
+fowler <- read.csv('modeling_data/fowler') %>%
    mutate(interestBee = str_to_title(interestBee)) %>%
@@
 
-globeeGenus <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteRawGenus') %>%
+globeeGenus <- read_csv('modeling_data/BeeCiteRawGenus') %>%
   dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
-globee_5cutGenus <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteC5Genus') %>%
+globee_5cutGenus <- read_csv('modeling_data/BeeCiteC5Genus') %>%
   dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
 
-globee_TGenus <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/BeeCiteWGenus') %>%
+globee_TGenus <- read_csv('modeling_data/BeeCiteWGenus') %>%
   dplyr::select(-west_specialist, -east_specialist, -cent_specialist, -specialist)
@@
 ```{r}
-interactions <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv')
+interactions <- read_csv('modeling_data/interactions_dup.csv')
 
@@
 
-write.csv(avery_send, "/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Family_Level_Transform.csv")
+# write.csv(avery_send, "/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Family_Level_Transform.csv")
 
@@
 final
-
-write.csv(final, '/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/visitation_specialization_df')
+write_csv(compare_df,'modeling_data/compare_df.csv')
+write_csv(final, 'modeling_data/visitation_specialization_df')
 ```

### 2022-07-21 scripts/compare ms and fowler data.R
changing path names

@@ -1,0 +1,31 @@
+rm(list = ls())
+library(tidyverse)
+select = dplyr::select; map = purrr::map; rename = dplyr::rename
+
+ms <- read_xlsx('modeling_data/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 3) %>%
+  rename_all(tolower) %>%
+  mutate(scientificName = paste0(genus, ' ',sub('.*\\. ','',species))) %>%
+  select(scientificName,everything()) %>%
+  rename(diet_breadth=`redefined pollen diet breadth`)
+
+fowler <- read_csv("modeling_data/fowler") %>%
+  mutate(scientificName = paste0(toupper(substr(interestBee,1,1)),
+                                 substr(fowler$interestBee,2,nchar(interestBee)))) %>%
+  select(scientificName,everything())
+
+
+#any bees in the fowler list not in the missouri state list?
+data.frame(fowler %>% filter(!scientificName %in% ms$scientificName))
+
+#any bees classified differently between the two list
+# ie classified as specialists by fowler but polylectic by missouri state
+data.frame(ms %>% 
+  filter(diet_breadth == 'Polylectic' & scientificName %in% fowler$scientificName))
+
+# (north american) bees on the missouri state list that are classified as specialists and 
+# not on jarrod fowlers list
+ms %>% 
+  filter(diet_breadth != "Polylectic" & !scientificName %in% fowler$scientificName) #the handful of these I checked don't apper to be north american
+#note some of the fowler bees have ** astericks near the name so that is something i need to deal with
+
+#to fi

### 2022-07-26 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep-reformatted.Rmd
starting script

--- 2022-07-21 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep.Rmd
@@
 knitr::opts_knit$set(error=T)
+select = dplyr::select
 
@@
 interactions_b <- read.csv('modeling_data/interactions_dup.csv')
+ms <- read_xlsx('modeling_data/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 3) %>%
+  rename_all(tolower) %>%
+  mutate(scientificName = paste0(genus, ' ',sub('.*\\. ','',species))) %>%
+  select(scientificName,everything()) %>%
+  rename(diet_breadth=`redefined pollen diet breadth`)
+fowler <- read_csv("modeling_data/fowler_hostplants.csv")
+
+#who are the specialsits?
+#let's say specialist the MS list and everything on fowler list that's missing from ms list
+ms_oligo = ms$scientificName[ms$diet_breadth %in% c('Oligolectic','Monolectic')]
+add_from_fowler = unique(fowler$scientificName[!fowler$scientificName %in% ms$scientificName])
+specialists = c(ms_oligo,add_from_fowler)
 
@@
 
-for (c in colnames(interactions_b)) {
-  interactions_b[[c]] <-  tolower(interactions_b[[c]])
-}
+# for (c in colnames(interactions_b)) {
+#   interactions_b[[c]] <-  tolower(interactions_b[[c]])
+# }
 ```
@@
 ```{r}
-interactions_b[1:5,]
+#how many species in the missouri state data are in the globi data?
+head(interactions_b)
+n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms$scientificName)]) #800 species in the ms data
+
+#how many are specialists
+ms_specialists = ms %>% filter(diet_breadth != 'Polylectic')
+n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_specialists$scientificName)]) #503 specialist species
+
+#how many are generalists
+ms_generalists = ms %>% filter(diet_breadth == 'Polylectic')
+n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_generalists$scientificName)]) #296 generalist species
+
+```
+
+```{r}
+interactions_b[1:5,1:5]
+```
+
+Let's do some reformatting
+```{r}
+capitalize = function(a_string){
+
+  beginning = toupper(substr(a_string,1,1))
+  ending = substr(a_string,2,nchar(a_string))
+
+  return(paste0(beginning,ending))
+
+}
+
+globi_summ = interactions_b %>% 
+  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
+         plant_family = capitalize(targetTaxonFamilyName), 
+         plant_genus = capitalize(targetTaxonGenusName)) %>%
+  filter(scientificName != '' & plant_family !='' & plant_genus != "") %>% 
+  group_by(scientificName, plant_genus, plant_family,) %>% 
+  summarize(n_genus = n()) %>% ungroup()
+
+```
+
+how many specialist bees do we have if we filter out interactions with <5 observations?
+```{r}
+globi_5n = globi_summ %>% filter(n_genus >= 5)
+n_distinct(globi_5n$scientificName[globi_5n$scientificName %in% ms_specialists$scientificName]) #only 360
+
+#how many other species
+n_distinct(globi_5n$scientificName[!globi_5n$scientificName %in% ms_specialists$scientificName]) # 730, so about 1000 total
+
+```
+Calculate number of interaction partners
+
+```{r}
+#calculate the number of plant families each bee species is observed visiting
+total_obs = globi_summ %>% group_by(scientificName) %>% summarize(n=sum(n_genus))
+
+# degree_by_fam 
+degree_by_fam = globi_summ %>% group_by(scientificName,plant_family) %>%
+  summarize(n_family = sum(n_genus)) %>%
+  left_join(total_obs) %>%
+  mutate(freq_squared = (n_family/n)^2) %>% group_by(scientificName,n) %>%
+  summarize(degree_family = n_distinct(plant_family), 
+            degree5_family = sum(n_family >= 5), weighted_degree_fam = sum(freq_squared))# %>%
+
+#calculate the number of plant genera each bee species is observed visiting
+degree_by_genus = globi_summ %>%
+  left_join(total_obs) %>%
+  mutate(freq_squared = (n_genus/n)^2) %>% 
+  group_by(scientificName) %>%
+  summarize(degree_genus = n_distinct(plant_genus), 
+            degree5_genus = sum(n_genus >= 5), weighted_degree_genus = sum(freq_squared))
+
+globi_degree = degree_by_fam %>% left_join(degree_by_genus) %>%
+  mutate(diet_breadth = ifelse(scientificName %in% specialists,'specialist','generalist'))
+
+write_csv(globi_degree,'modeling_data/globi_degree.csv')
+
 ```
@@
 table_test$specialist[is.na(table_test$specialist)] <- 0
+getwd()
 
-write_csv(table_test, 'modeling_data/fowler')
+# write_csv(table_test, 'modeling_data/fowler')
+```
+```{r reformat the table to include host plant data}
+fowler_table = eastern_table %>% mutate(region = 'east') %>%
+  bind_rows(central_table %>% mutate(region='central')) %>%
+  bind_rows(western_table %>% mutate(region = 'west')) %>%
+  rename(host_plant = Host.plant.Family..Tribe..Genera) %>%
+  rename_all(tolower) %>%
+  mutate(scientificName = gsub(" *\\(.*?\\) *", "", genus..subgenus..species)) %>%
+  select(scientificName,authority,status,host_plant,region) %>%
+  #remove astericks from some of the scientific names:
+  mutate(scientificName = ifelse(grepl("\\*",scientificName),sub("\\*.*","",scientificName),scientificName))
+
+# write_csv(fowler_table,'modeling_data/fowler_hostplants.csv')
 ```
@@
 plotdf <- merge(beeCite, table_test, by = 'interestBee', all.x = TRUE)
+
+nrow(plotdf)
+
 plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist

### 2022-07-26 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep-reformatted.Rmd
deleting old code

@@
 
-write_csv(globi_degree,'modeling_data/globi_degree.csv')
+# write_csv(globi_degree,'modeling_data/globi_degree.csv')
 
 ```
-
-Define beeCites with varying criteria
-
-```{r, fig.width=5,fig.height=2}
-beeCiteNoFilter <- interactions_b %>% 
-  dplyr::select(sourceTaxonSpeciesName,targetTaxonFamilyName) %>% 
-  dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonFamilyName) %>% dplyr::group_by(interestBee, interact) %>% dplyr::count() %>% ungroup() %>% 
-  filter(interestBee != "", interact != "")
-
-TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% dplyr::summarise(n_Bee = sum(n)) %>% ungroup()
-TotalCitations <- sum(TotalCitations$n_Bee)
-
-beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% dplyr::summarise(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
-beeCiteNoFilter <- beeCiteNoFilter %>% dplyr::summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup() %>% unique()
-
-beeIntDegree <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count() 
-beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique()
-
-beeIntDegreeC10 <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
-beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
-beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
-beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 10 citations to contribute to degree
-
-beeIntDegreeC5 <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
-beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
-beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
-beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 5 citations to contribute to degree
-
-beeIntDegreeW <- beeCiteNoFilter %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
-beeIntDegreeW$numcitations <- beeCiteNoFilter$n
-beeIntDegreeW$n_Bee <- beeCiteNoFilter$n_Bee
-beeIntDegreeW$n <- (beeIntDegreeW$numcitations / beeIntDegreeW$n_Bee)^2
-beeIntDegreeW <- beeIntDegreeW %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## weighted degree
-
-beeCite <-  left_join(beeCiteNoFilter, beeIntDegree, by = 'interestBee')
-
-beeCiteC10 <-  left_join(beeCiteNoFilter, beeIntDegreeC10, by = 'interestBee')
-
-beeCiteC5 <-  left_join(beeCiteNoFilter, beeIntDegreeC5, by = 'interestBee')
-
-beeCiteW <-  left_join(beeCiteNoFilter, beeIntDegreeW, by = 'interestBee')
-
-beeCiteNoFilterGenus <- interactions_b %>% dplyr::select(sourceTaxonSpeciesName, targetTaxonGenusName) %>% 
-  dplyr::rename(interestBee = sourceTaxonSpeciesName, interact = targetTaxonGenusName) %>% group_by(interestBee, interact) %>% dplyr::count() %>% ungroup() %>% 
-  filter(interestBee != "", interact != "")
-
-TotalCitationsGenus <- beeCiteNoFilterGenus %>% group_by(interestBee) %>% dplyr::summarise(n_Bee = sum(n)) %>% ungroup()
-TotalCitationsGenus <- sum(TotalCitationsGenus$n_Bee)
-
-beeCiteNoFilterGenus <- beeCiteNoFilterGenus %>% group_by(interestBee) %>% dplyr::summarise(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
-beeCiteNoFilterGenus <- beeCiteNoFilterGenus %>% dplyr::summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()
-
-beeIntDegreeGenus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
-beeIntDegreeGenus <- beeIntDegreeGenus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique()
-
-beeIntDegreeC10Genus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
-beeIntDegreeC10Genus$numcitations <- beeCiteNoFilterGenus$n
-beeIntDegreeC10Genus <- beeIntDegreeC10Genus %>% filter(numcitations >= 10)
-beeIntDegreeC10Genus <- beeIntDegreeC10Genus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 10 citations to contribute to degree
-
-beeIntDegreeC5Genus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
-beeIntDegreeC5Genus$numcitations <- beeCiteNoFilterGenus$n
-beeIntDegreeC5Genus <- beeIntDegreeC5Genus %>% filter(numcitations >= 5)
-beeIntDegreeC5Genus <- beeIntDegreeC5Genus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## min of 5 citations to contribute to degree
-
-beeIntDegreeWGenus <- beeCiteNoFilterGenus %>% dplyr::select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% dplyr::count()
-beeIntDegreeWGenus$numcitations <- beeCiteNoFilterGenus$n
-beeIntDegreeWGenus$n_Bee <- beeCiteNoFilterGenus$n_Bee
-beeIntDegreeWGenus$n <- (beeIntDegreeWGenus$numcitations / beeIntDegreeWGenus$n_Bee)^2
-beeIntDegreeWGenus <- beeIntDegreeWGenus %>% group_by(interestBee) %>% dplyr::summarise(interestBee = interestBee, degree = sum(n)) %>% unique() ## weighted degree
-
-beeCiteGenus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeGenus, by = 'interestBee')
-
-beeCiteC10Genus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeC10Genus, by = 'interestBee')
-
-beeCiteC5Genus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeC5Genus, by = 'interestBee')
-
-beeCiteWGenus <-  left_join(beeCiteNoFilterGenus, beeIntDegreeWGenus, by = 'interestBee')
-
-beeCite[1:10,]
-```
-
-```{r}
-beeCiteGenus <- beeCiteGenus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
-                        bee_proportion_of_all_bee_citations = BeeFreq, degreeGenus = degree)
-
-beeCite <- beeCite %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
-                        bee_proportion_of_all_bee_citations = BeeFreq, degreeFamily = degree)
-
-beeCiteC5Genus <- beeCiteC5Genus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
-                        bee_proportion_of_all_bee_citations = BeeFreq, degreeC5Genus = degree)
-
-beeCiteC5 <- beeCiteC5 %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
-                        bee_proportion_of_all_bee_citations = BeeFreq, degreeC5Family = degree)
-
-beeCiteWGenus <- beeCiteWGenus %>% dplyr::rename(interactGenus = interact, number_of_citations_of_genus_interaction = n, number_of_citation_for_bee_species = n_Bee, 
-                        bee_proportion_of_all_bee_citations = BeeFreq, degreeWGenus = degree)
-beeCiteW <- beeCiteW %>% dplyr::rename(interactFamily = interact, number_of_citations_of_family_interaction = n, number_of_citation_for_bee_species = n_Bee, 
-                        bee_proportion_of_all_bee_citations = BeeFreq, degreeWFamily = degree)
-
-beeCiteW[1:5,]
-```
-
-## Include Fowler Data
-
-```{r}
-western_page <- read_html("https://jarrodfowler.com/pollen_specialist.html")
-eastern_page <- read_html("https://jarrodfowler.com/specialist_bees.html")
-central_page <- read_html("https://jarrodfowler.com/bees_pollen.html")
-
-western_table <- western_page %>% html_table(fill=TRUE)
-eastern_table<- eastern_page %>% html_table(fill=TRUE)
-central_table<- central_page %>% html_table(fill=TRUE)
-
-western_table <- as.data.frame(western_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
-eastern_table <- as.data.frame(eastern_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
-central_table <- as.data.frame(central_table) %>% dplyr::select(-Family..Subfamily..Tribe..Subtribe)
-
-western_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", western_table$Genus..Subgenus..species))
-eastern_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", eastern_table$Genus..Subgenus..species))
-central_table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", central_table$Genus..Subgenus..species))
-
-table <- rbind(western_table, eastern_table, central_table)
-table$interestBee <- tolower(gsub(" *\\(.*?\\) *", "", table$Genus..Subgenus..species))
-
-western_table$west_specialist <- 1
-western_test <- western_table %>% dplyr::select(interestBee, west_specialist, Status) %>% dplyr::rename('Status_W' = 'Status') %>% unique()
-
-eastern_table$east_specialist <- 1
-eastern_test <- eastern_table %>% dplyr::select(interestBee, east_specialist, Status)%>% dplyr::rename('Status_E'= 'Status') %>% unique()
-
-central_table$cent_specialist <- 1
-central_test <- central_table %>% dplyr::select(interestBee, cent_specialist, Status)%>% dplyr::rename('Status_C'= 'Status') %>% unique()
-
-tablejoining <- merge(x = western_test, y = eastern_test, by = "interestBee", all = TRUE)
-table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)
-
-table_test$specialist <- table_test$west_specialist + table_test$east_specialist + table_test$cent_specialist
-table_test$specialist <- as.integer(table_test$specialist > 0)
-table_test$specialist[is.na(table_test$specialist)] <- 0
-getwd()
-
-# write_csv(table_test, 'modeling_data/fowler')
-```
-```{r reformat the table to include host plant data}
-fowler_table = eastern_table %>% mutate(region = 'east') %>%
-  bind_rows(central_table %>% mutate(region='central')) %>%
-  bind_rows(western_table %>% mutate(region = 'west')) %>%
-  rename(host_plant = Host.plant.Family..Tribe..Genera) %>%
-  rename_all(tolower) %>%
-  mutate(scientificName = gsub(" *\\(.*?\\) *", "", genus..subgenus..species)) %>%
-  select(scientificName,authority,status,host_plant,region) %>%
-  #remove astericks from some of the scientific names:
-  mutate(scientificName = ifelse(grepl("\\*",scientificName),sub("\\*.*","",scientificName),scientificName))
-
-# write_csv(fowler_table,'modeling_data/fowler_hostplants.csv')
-```
-
-```{r}
-plotdf <- merge(beeCite, table_test, by = 'interestBee', all.x = TRUE)
-
-nrow(plotdf)
-
-plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
-plotdf$specialist <- as.integer(plotdf$specialist > 0)
-plotdf$specialist[is.na(plotdf$specialist)] <- 0
-
-write_csv(plotdf, 'modeling_data/BeeCiteRaw')
-```
-
-```{r}
-plotdf <- merge(beeCiteC5, table_test, by = 'interestBee', all.x = TRUE)
-plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
-plotdf$specialist <- as.integer(plotdf$specialist > 0)
-plotdf$specialist[is.na(plotdf$specialist)] <- 0
-
-write_csv(plotdf, 'modeling_data/BeeCiteC5')
-```
-
-```{r}
-plotdf <- merge(beeCiteW, table_test, by = 'interestBee', all.x = TRUE)
-plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
-plotdf$specialist <- as.integer(plotdf$specialist > 0)
-plotdf$specialist[is.na(plotdf$specialist)] <- 0
-
-write_csv(plotdf, 'modeling_data/BeeCiteW')
-```
-
-```{r}
-plotdf <- merge(beeCiteGenus, table_test, by = 'interestBee', all.x = TRUE)
-plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
-plotdf$specialist <- as.integer(plotdf$specialist > 0)
-plotdf$specialist[is.na(plotdf$specialist)] <- 0
-
-write_csv(plotdf, 'modeling_data/BeeCiteRawGenus')
-```
-
-```{r}
-plotdf <- merge(beeCiteC5Genus, table_test, by = 'interestBee', all.x = TRUE)
-plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
-plotdf$specialist <- as.integer(plotdf$specialist > 0)
-plotdf$specialist[is.na(plotdf$specialist)] <- 0
-
-write_csv(plotdf, 'modeling_data/BeeCiteC5Genus')
-```
-
-```{r}
-plotdf <- merge(beeCiteWGenus, table_test, by = 'interestBee', all.x = TRUE)
-plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
-plotdf$specialist <- as.integer(plotdf$specialist > 0)
-plotdf$specialist[is.na(plotdf$specialist)] <- 0
-
-write_csv(plotdf, 'modeling_data/BeeCiteWGenus')
-```

### 2022-07-27 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep-reformatted.Rmd
update plant names using the plant list

@@
 library(data.table)
+library(readxl)
+library(Taxonstand)
 
@@
 
-Let's do some reformatting
+Let's reformat and update the plant names in the interaction data 
 ```{r}
@@
 
-globi_summ = interactions_b %>% 
+interactions_reformat = interactions_b %>% 
   mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
+         bee_family = capitalize(sourceTaxonFamilyName),
          plant_family = capitalize(targetTaxonFamilyName), 
-         plant_genus = capitalize(targetTaxonGenusName)) %>%
-  filter(scientificName != '' & plant_family !='' & plant_genus != "") %>% 
-  group_by(scientificName, plant_genus, plant_family,) %>% 
-  summarize(n_genus = n()) %>% ungroup()
+         plant_genus = capitalize(targetTaxonGenusName),
+         plant_species = capitalize(targetTaxonName)) %>%
+  filter(scientificName != '' & plant_family !='' & plant_genus != "")
+
+#
+
+plant_df = interactions_reformat %>%
+  distinct(plant_species)
+test_me = plant_df$plant_species
+
+# check_names = TPL(test_me) #takes forever to run, download output as csv
+# write_csv(check_names,'modeling_data/plant_list_name_update.csv')
+check_names = read_csv('modeling_data/plant_list_name_update.csv')
+
+check_names %>% filter(Taxonomic.status != 'Accepted') %>% 
+  select(Taxon,New.Genus,New.Species)
+
+tpl_formatted = check_names %>% rename(plant_species = Taxon) %>% 
+  mutate(accepted_name = ifelse(!is.na(New.Species),paste(New.Genus,New.Species),New.Genus) ) %>%
+  mutate(accepted_name = ifelse(is.na(accepted_name), plant_species, accepted_name)) %>%
+  select(plant_species,accepted_name)
+
+tpl_formatted %>% filter(is.na(accepted_name))
+
+#add accepted_name column to plant_df
+interactions_updated = interactions_reformat %>% 
+  group_by(scientificName,bee_family,plant_species,plant_family) %>%
+  summarize(n=n()) %>%
+  ungroup %>%
+  left_join(tpl_formatted) %>%
+  select(-plant_species) %>%
+  rename(plant_species = accepted_name) %>%
+  mutate(plant_genus = sub(" .*","",plant_species))
+
+```
+
+Let's do some reformatting
+```{r}
+
+globi_summ = interactions_updated %>%
+  group_by(scientificName, plant_genus, plant_family,bee_family) %>%
+  summarize(n_genus = sum(n)) %>% ungroup()
 
@@
   mutate(freq_squared = (n_genus/n)^2) %>% 
-  group_by(scientificName) %>%
+  group_by(scientificName,bee_family) %>%
   summarize(degree_genus = n_distinct(plant_genus), 
@@
   mutate(diet_breadth = ifelse(scientificName %in% specialists,'specialist','generalist'))
-
-# write_csv(globi_degree,'modeling_data/globi_degree.csv')
+getwd()
+# write_csv(globi_degree,'modelign_data/globi_degree.csv')
 

### 2022-10-13 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep-reformatted.Rmd
removing interactions not at species or genus level

@@
 library(Taxonstand)
+library(vroom)
 
@@
 options(max.print=1000000)
-interactions_b <- read.csv('modeling_data/interactions_dup.csv')
-ms <- read_xlsx('modeling_data/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 3) %>%
-  rename_all(tolower) %>%
-  mutate(scientificName = paste0(genus, ' ',sub('.*\\. ','',species))) %>%
-  select(scientificName,everything()) %>%
-  rename(diet_breadth=`redefined pollen diet breadth`)
+interactions_b <- vroom('modeling_data/interactions_dup.csv')
 fowler <- read_csv("modeling_data/fowler_hostplants.csv")
-
-#who are the specialsits?
-#let's say specialist the MS list and everything on fowler list that's missing from ms list
-ms_oligo = ms$scientificName[ms$diet_breadth %in% c('Oligolectic','Monolectic')]
-add_from_fowler = unique(fowler$scientificName[!fowler$scientificName %in% ms$scientificName])
-specialists = c(ms_oligo,add_from_fowler)
 
@@
 }
-
-# for (c in colnames(interactions_b)) {
-#   interactions_b[[c]] <-  tolower(interactions_b[[c]])
-# }
-```
-
-```{r}
-#how many species in the missouri state data are in the globi data?
-head(interactions_b)
-n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms$scientificName)]) #800 species in the ms data
-
-#how many are specialists
-ms_specialists = ms %>% filter(diet_breadth != 'Polylectic')
-n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_specialists$scientificName)]) #503 specialist species
-
-#how many are generalists
-ms_generalists = ms %>% filter(diet_breadth == 'Polylectic')
-n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_generalists$scientificName)]) #296 generalist species
 
@@
          plant_species = capitalize(targetTaxonName)) %>%
-  filter(scientificName != '' & plant_family !='' & plant_genus != "")
+  filter(scientificName != '' & plant_family !='' & plant_genus != "")%>%
+   filter(!targetTaxonRank %in% c('species',"genus"))
 
 #
+str(interactions_reformat)
 
@@
 
-tpl_formatted = check_names %>% rename(plant_species = Taxon) %>% 
+tpl_formatted = check_names %>% 
+  rename(plant_species = Taxon) %>% 
   mutate(accepted_name = ifelse(!is.na(New.Species),paste(New.Genus,New.Species),New.Genus) ) %>%
@@
   group_by(scientificName,bee_family,plant_species,plant_family) %>%
-  summarize(n=n()) %>%
+  summarize(n=n(),n_source_citations = n_distinct(sourceCitation),n_ref_citations = n_distinct(referenceCitation)) %>%
   ungroup %>%
@@
   rename(plant_species = accepted_name) %>%
-  mutate(plant_genus = sub(" .*","",plant_species))
+  mutate(plant_genus = sub(" .*","",plant_species))%>%
+  mutate(plant_genus = ifelse(plant_genus == "Ipomaea","Ipomoea",plant_genus))
 
+head(interactions_updated)
+
+#let's also reformat so that the data re not grouped (each row a unique record)
+nrow(interactions_reformat %>% left_join(tpl_formatted)) == nrow(interactions_reformat)
+
+str(interactions_reformat)
+globi_names = colnames(interactions_reformat)
+globi_names[grepl('Sex',globi_names)]
+
+each_interaction = interactions_reformat %>% 
+  select(scientificName,bee_family,plant_species,plant_family,interactionTypeName,sourceTaxonRank, targetTaxonRank,referenceCitation,sourceCitation,decimalLongitude,decimalLatitude,sourceSexId,sourceSexName,targetSexId,targetSexName) %>%
+  left_join(tpl_formatted) %>%
+  select(-plant_species) %>%
+  rename(plant_species = accepted_name) %>%
+  mutate(plant_genus = sub(" .*","",plant_species))%>%
+  mutate(plant_genus = ifelse(plant_genus == "Ipomaea","Ipomoea",plant_genus))
+
+interactions_updated %>%
+  select(scientificName,n,n_source_citations,n_ref_citations,plant_species) %>%
+  filter(n>5)
+
+# write_csv(interactions_reformat %>% filter(scientificName == "Agapostemon angelicus" & plant_species =="Helianthus annuus") %>% distinct(sourceCitation,referenceCitation),'look_at_refs.csv')
+
+# write_csv(interactions_updated,'modeling_data/globi_names_updated.csv')
+# write_csv(each_interaction,'modeling_data/globi_occ_names_updated.csv')
 ```
@@
   group_by(scientificName, plant_genus, plant_family,bee_family) %>%
-  summarize(n_genus = sum(n)) %>% ungroup()
+  summarize(n_genus = sum(n),genus_source_citations=sum(n_source_citations)) %>% #this treats interactions of different plant species within the same genus as separate citations
+  ungroup()
 
@@
 ```{r}
+
 #calculate the number of plant families each bee species is observed visiting
@@
 degree_by_fam = globi_summ %>% group_by(scientificName,plant_family) %>%
-  summarize(n_family = sum(n_genus)) %>%
+  summarize(n_family = sum(n_genus),n_source_citations=sum(genus_source_citations)) %>% # %>%this treats genera within the same fam, from the same source as separate sources
   left_join(total_obs) %>%
-  mutate(freq_squared = (n_family/n)^2) %>% group_by(scientificName,n) %>%
+  mutate(freq_squared = (n_family/n)^2) %>% 
+  group_by(scientificName,n) #%>%
   summarize(degree_family = n_distinct(plant_family), 
-            degree5_family = sum(n_family >= 5), weighted_degree_fam = sum(freq_squared))# %>%
+            degree5_family = sum(n_family >= 5), weighted_degree_fam = sum(freq_squared),
+            simpson_family_citations=sum(n_source_citations),simpson())
 
@@
   mutate(diet_breadth = ifelse(scientificName %in% specialists,'specialist','generalist'))
-getwd()
 # write_csv(globi_degree,'modelign_data/globi_degree.csv')

### 2022-12-15 scripts/update plant names.R
starting script

@@ -1,0 +1,62 @@
+
+#update plant names for both globi data and fowler data using catalog of life
+library(vroom)
+library(tidyverse)
+
+#load the data
+options(max.print=1000000)
+interactions_b <- vroom('modeling_data/interactions-14dec2022.csv')
+fowler <- read_csv("modeling_data/fowler_hostplants.csv") 
+
+for (c in colnames(interactions_b)) {
+  interactions_b[[c]] <-  tolower(interactions_b[[c]])
+}
+
+
+capitalize = function(a_string){
+  
+  beginning = toupper(substr(a_string,1,1))
+  ending = substr(a_string,2,nchar(a_string))
+  
+  return(paste0(beginning,ending))
+  
+}
+
+interactions_reformat = interactions_b %>% 
+  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
+         bee_family = capitalize(sourceTaxonFamilyName),
+         plant_family = capitalize(targetTaxonFamilyName), 
+         plant_genus = capitalize(targetTaxonGenusName),
+         plant_species = capitalize(targetTaxonName)) %>%
+  filter(scientificName != '' & plant_family !='' & plant_genus != "") %>%
+  filter(targetTaxonRank %in% c('subspecies','species',"genus",'subgenus','variety')) %>%
+  filter(sourceTaxonRank %in% c("subspecies",'variety','species')) %>%
+  mutate(plant_family = ifelse(plant_family=='Compositae','Asteraceae',plant_family)) %>%
+  mutate(plant_family = ifelse(plant_family=='Umbelliferae','Apiaceae',plant_family)) %>%
+  mutate(plant_family = ifelse(plant_family=='Labiatae','Lamiaceae',plant_family)) 
+
+plant_df = interactions_reformat %>%
+  distinct(plant_species)
+test_me = plant_df$plant_species
+
+
+#load world flora online data
+wfo_data = vroom("/Volumes/Seagate/globi_13dec2022/data/WFO_Backbone/classification.txt") %>%
+  mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) # add a field with a scientific name for subspecies
+
+#filter the wfo data to just be species that are in the globi data
+wfo_filtered = wfo_data %>% filter(scientificName %in% test_me | scientificName2 %in% test_me)
+
+#look at the taxa that are not in the data:
[…]

### 2022-12-15 scripts/main analyses.R
starting script

@@ -373,15 +373,30 @@ globi_degree125 %>%
 #run the random forest model
 set.seed(20)
 rf_all = randomForest(as.factor(diet_breadth) ~ degree_family + degree_genus + simpson_fam + simpson_genus + bee_family + n ,
-                      data = globi_degree125,importance = T)
-set.seed(20)
-rf_all2 = randomForest(as.factor(diet_breadth) ~ degree_family + degree_genus + simpson_fam + simpson_genus + bee_family + n ,
-                     data = globi_degree125,importance = T)
+                      data = globi_degree125,importance = T,nperm=1)
+species_probs = data.frame(predict(rf_all,type='prob'))
+
+#any specialists on our list that have low probability of being specialists?
+head(globi_degree125)
+prob_output=globi_degree125 %>% bind_cols(species_probs)
+with(prob_output,boxplot(generalist~diet_breadth))
+
+data.frame(prob_output %>% 
+             filter(diet_breadth=='specialist'&specialist<.5) %>%
+             select(scientificName,simpson_fam,specialist))
+
+#let's see if species with <0.5 prob of being specialists have higher simpson div of plant fams visited
+with(prob_output %>%
+  filter(diet_breadth =='specialist') %>%
+  mutate(low_prob = specialist<0.5),boxplot(low_prob,simpson_fam))
+with(prob_output %>% filter(diet_breadth=='specialist'),
+     plot(specialist,simpson_fam))
+head(globi_degree125)
+View(globi_degree125)
 
 rf_w_citations = randomForest(as.factor(diet_breadth) ~ degree_family + degree_genus + simpson_fam + simpson_genus + bee_family + n + simpson_citation_genus+simpson_citation_fam,data = globi_degree125,importance = T)
 
 rf_all
-rf_all2
 importance(rf_all)
 rf_all$confusion
 globi_degree %>% filter(scientificName == 'Andrena erigeneae')
@@ -417,27 +432,38 @@ varImpPlot(rf_all)
 # dev.off()
 
 imp = data.frame(rf_all$importance)
+imp = data.frame(importance(rf_all)) 
 imp$predictor = rownames(imp)
+imp$labels = c('plant families visited',"plant genera visited","plant families visited \n(simpson)",
+               "plant genera visited \n(simpson)","bee family","sample size")
[…]

### 2022-12-16 scripts/update plant names.R
update synonyms

@@ -1,4 +1,4 @@
-
+rm(list=ls())
 #update plant names for both globi data and fowler data using catalog of life
 library(vroom)
 library(tidyverse)
@@ -22,6 +22,7 @@ capitalize = function(a_string){
   
 }
 
+#reformat the interactions data:
 interactions_reformat = interactions_b %>% 
   mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
          bee_family = capitalize(sourceTaxonFamilyName),
@@ -39,24 +40,179 @@ plant_df = interactions_reformat %>%
   distinct(plant_species)
 test_me = plant_df$plant_species
 
+# #load world flora online data
+# wfo_data = vroom("/Volumes/Seagate/globi_13dec2022/data/WFO_Backbone/classification.txt") %>%
+#   mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) %>% # add a field with a scientific name for subspecies
+#   mutate(scientificName3 = paste(genus,specificEpithet))
 
 #load world flora online data
-wfo_data = vroom("/Volumes/Seagate/globi_13dec2022/data/WFO_Backbone/classification.txt") %>%
-  mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) # add a field with a scientific name for subspecies
+wfo_data = vroom("modeling_data/WFO_Backbone/classification.txt") %>%
+  mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) %>% # add a field with a scientific name for subspecies
+  mutate(scientificName3 = paste(genus,specificEpithet))
 
-#filter the wfo data to just be species that are in the globi data
-wfo_filtered = wfo_data %>% filter(scientificName %in% test_me | scientificName2 %in% test_me)
 
+#these species were not in wfo, and I looked them up manually in catalog of life or itis:
[…]

### 2022-12-16 modeling_data/GLOBI_bee_update_clean.Rmd
update to write_csv
switching from data.table to vroom

@@
 ```{r}
-library(data.table)
+#library(data.table)
+#df <- fread( 'https://zenodo.org/record/4460654/files/interactions.csv.gz')
+library(vroom)
+library(tidyverse)
 
-df <- fread( 'https://zenodo.org/record/4460654/files/interactions.csv.gz')
+df = vroom("all_bee_data.tsv")
+
 ```
@@
 # Flip target/source for above df
-
 colnames(plant_source_df) <- as.character(lapply(colnames(plant_source_df), textclean::swap, pattern1 = 'source', pattern2 = 'target'))
-
 plant_source_df <- plant_source_df %>% dplyr::rename(sourceCitation = targetCitation, sourceNamespace = targetNamespace, 
@@
 interactions_dup = rbind(bee_source_df, plant_source_df) 
-
 # Combine and drop duplicates
@@
 ```{r}
-write.csv(interactions_dup, "interactions_dup.csv")
+readr::write_csv(interactions, "interactions-14dec2022.csv")
 ```

### 2022-12-19 scripts/update plant names.R
adding line for saving final output

@@ -40,15 +40,15 @@ plant_df = interactions_reformat %>%
   distinct(plant_species)
 test_me = plant_df$plant_species
 
-# #load world flora online data
-# wfo_data = vroom("/Volumes/Seagate/globi_13dec2022/data/WFO_Backbone/classification.txt") %>%
-#   mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) %>% # add a field with a scientific name for subspecies
-#   mutate(scientificName3 = paste(genus,specificEpithet))
-
 #load world flora online data
-wfo_data = vroom("modeling_data/WFO_Backbone/classification.txt") %>%
+wfo_data = vroom("/Volumes/Seagate/globi_13dec2022/data/WFO_Backbone/classification.txt") %>%
   mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) %>% # add a field with a scientific name for subspecies
   mutate(scientificName3 = paste(genus,specificEpithet))
+
+# #load world flora online data
+# wfo_data = vroom("modeling_data/WFO_Backbone/classification.txt") %>%
+#   mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) %>% # add a field with a scientific name for subspecies
+#   mutate(scientificName3 = paste(genus,specificEpithet))
 
 
 #these species were not in wfo, and I looked them up manually in catalog of life or itis:
@@ -170,9 +170,7 @@ wfo_unsure %>% filter(old_name == "Taraxacum vulgare")
 # 5) for all synonyms:
 #   filter wfo_filtered to just be these species (filter so  that none are also accepted)
 
-wfo_filtered2 %>%
-  filter(scientificName %in% not_accepted) %>%
-  distinct(taxonomicStatus)
+
 # View(wfo_filtered2 %>% filter(taxonomicStatus=="SYNONYM")) #doubtful and unchecked don't have accepted usage-ids, so we'll just go with those
 # View(wfo_filtered2 %>% filter(taxonID=="wfo-4000035688"))
 # View(wfo_filtered2 %>% filter(taxonomicStatus=="SYNONYM" & is.na(acceptedNameUsageID)))
@@ -198,8 +196,8 @@ wfo_name_update = wfo_accepted %>% bind_rows(wfo_syn_update) %>% bind_rows(wfo_u
   rename(old_plant = old_name)
 
 #double check all plants are in the df and that there are no duplicates
[…]

### 2022-12-19 scripts/main analyses.R
updated plant names in fowler with world flora online

@@ -25,52 +25,15 @@ pollen_fidelity = read_csv("modeling_data/prop_host_pollen.csv")
 
 # load the globi data
 # this is a file with the globi records and plant names are updated (though not plant families)
-globi_r = read_csv('modeling_data/globi_occ_names_updated.csv')
+globi_r = read_csv('modeling_data/globi_occ_names_updated-19dec2022.csv')
 globi_r %>% filter(sourceTaxonRank=='variety') %>% distinct(scientificName)
 data.frame(globi_r %>% filter(is.na(sourceTaxonRank)) %>% distinct(scientificName))
 
 
 #also load the fowler data
-#some of these bees have different characterizations despite having the same pollen hosts in both databases
-# this is because fowler considers bees to be specialists if they use pollen from genera in two different families
-#and we don't. we're going to remove these bees from the fowler list (and won't consider them discrepancies)
-generalists_fowler = c("Andrena candidiformis", "Anthidium mormonum", "Dufourea cuprea",
-                       "Habropoda laboriosa", "Hesperapis ilicifoliae", "Megachile perihirta",
-                       "Peponapis michelbacherorum",
-                       "Perdita fieldi","Perdita obscurata",
-                       "Pseudopanurgus virginicus", "Xenoglossa kansensis","Florilegus condignus")
-
-#note: some bees on this list are duplicated, if they occur in multiple regions (eg both eastern and central usa)
-fowler <- read_csv("modeling_data/fowler_hostplants.csv") %>%
-  mutate(diet_breadth = ifelse(scientificName %in% generalists_fowler,'generalist','specialist')) %>%
-  mutate(host_plant_rank = ifelse(grepl('aceae',host_plant),'family','genus')) %>%
-  mutate(diet_breadth_detailed = ifelse(host_plant_rank == 'family' & diet_breadth =='specialist','family_specialist','genus_specialist')) %>%
-  mutate(diet_breadth_detailed = ifelse(diet_breadth=='generalist','generalist',diet_breadth_detailed))
-
-specialists = fowler[fowler$diet_breadth=='specialist',]$scientificName
-fowler_formatted = read_csv('modeling_data/fowler_formatted.csv')
-
-##
-# View(globi_r %>% distinct(sourceCitation,referenceCitation))
-
-
[…]

### 2022-12-19 scripts/specialists visiting non-hosts.R

@@ -1,7 +1,8 @@
 rm(list = ls())
 library(tidyverse)
 library(readxl)
-
+library(kewr)
+library(furrr)
 
 # Research Qs this R script will answer:
 # What percentage of oligolectic bees visit non-host plants (where visit == at least 5 visits)
@@ -86,12 +87,19 @@ with(prop_nonhost,plot(prop_nonhost_visits~total_n))
 
 with(prop_nonhost,boxplot(prop_nonhost_visits~only_ms))
 
+# format the box-plot graph:
+# pdf('figures/compare_fowler_ms.pdf')
+with(prop_nonhost %>% mutate(on_fowler = factor(!only_ms,levels = c(T,F))),
+     boxplot(prop_nonhost_visits~on_fowler, ylab = 'proportion of bees visiting non-hosts',
+                          xlab = 'On Fowler list'))
+# dev.off()
 
-#is there a difference between bees on fowler list vs MS list?
+
+# is there a difference between bees on fowler list vs MS list?
 only_ms_specialist$scientificName
 
-##old:
-#note: for fowler df, bees on this list are duplicated if they occur in multiple regions (eg both east and central america)
+## old:
+# note: for fowler df, bees on this list are duplicated if they occur in multiple regions (eg both east and central america)
 fowler <- read_csv("modeling_data/fowler_hostplants.csv") 
 globi <- read_csv('modeling_data/globi_names_updated.csv')
 

### 2022-12-19 scripts/format_fowler_hosts.R
starting script

@@ -1,0 +1,252 @@
+rm(list=ls())
+library(tidyverse)
+library(furrr)
+
+#load world flora online data for getting accepted species names
+wfo_data = vroom("/Volumes/Seagate/globi_13dec2022/data/WFO_Backbone/classification.txt") %>%
+  mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) %>% # add a field with a scientific name for subspecies
+  mutate(scientificName3 = paste(genus,specificEpithet))
+
+# #load world flora online data - from computer instead of seagate
+# wfo_data = vroom("modeling_data/WFO_Backbone/classification.txt") %>%
+#   mutate(scientificName2 = paste(genus,specificEpithet,infraspecificEpithet)) %>% # add a field with a scientific name for subspecies
+#   mutate(scientificName3 = paste(genus,specificEpithet))
+
+#these genus is not in wfo, and I looked it up manually in catalog of life :
+not_in_wfo = data.frame(fowler = c('Salazaria'),
+                        accepted = c("Scutellaria"),
+                        family = c("Lamiaceae"),
+                        source = c('col'))
+
+# fowler considers bees to be specialists if they use pollen from genera in two different families
+#and we don't. we're going to remove these bees from the fowler list (and won't consider them discrepancies)
+generalists_fowler = c("Andrena candidiformis", "Anthidium mormonum", "Dufourea cuprea",
+                       "Habropoda laboriosa", "Hesperapis ilicifoliae", "Megachile perihirta",
+                       "Peponapis michelbacherorum",
+                       "Perdita fieldi","Perdita obscurata",
+                       "Pseudopanurgus virginicus", "Xenoglossa kansensis","Florilegus condignus")
+
+#note: some bees on this list are duplicated, if they occur in multiple regions (eg both eastern and central usa)
+fowler <- read_csv("modeling_data/fowler_hostplants.csv") %>%
+  mutate(diet_breadth = ifelse(scientificName %in% generalists_fowler,'generalist','specialist')) %>%
+  mutate(host_plant_rank = ifelse(grepl('aceae',host_plant) | grepl('ieae',host_plant),'family','genus')) %>%
[…]

### 2022-12-19 scripts/format_fowler_hosts.R
deleting old code

@@ -192,61 +192,3 @@ nrow(fowler_formatted) == nrow(hosts_long)
 write_csv(fowler_formatted,'modeling_data/fowler_formatted.csv')
 
 
-##old
-# fowler_genus_update = Taxonstand::TPL(update_these_names)
-# fowler_genus_update %>% select(Genus,New.Genus) %>% filter(Genus != New.Genus) #no changes
-
-#get families for genera 
-# update the plant family names using world checklist for vascular plants
-plan(multisession,workers=6)
-get_fams = update_these_names %>% future_map(function(sciname){
-  
-  output = search_wcvp(sciname, filters=c("families", "accepted"),limit = 1)
-  output_df = tidy(output)
-
-  if(nrow(output_df) !=0){
-    return_df = data.frame(new_host = sciname,family = output_df$family)
-  }else{
-    output = search_wcvp(sciname, filters=c("families"))
-    df=tidy(output)
-
-    if(nrow(df) !=0 & n_distinct(df$family)==1){
-      new_output_df = df$synonymOf[[1]]
-      return_df = data.frame(new_host = sciname,family = new_output_df$family)
-
-    }
-    if(nrow(df) !=0 & n_distinct(df$family)!=1){
-      df_sum = df %>% group_by(family) %>% summarize(n=n()) %>% arrange(desc(n))
-      return_df = data.frame(new_host = sciname,family = df_sum$family[1])
-      
-      }
-    if(nrow(df)==0){
-      return_df = data.frame(new_host=sciname,family = NA)
-    }
-  }
-  return(return_df)
-
-  }) %>% bind_rows
-get_fams %>% filter(new_host == "Horkelia")
-
-fowler_formatted = hosts_long %>% left_join(get_fams) %>%
-  mutate(family = ifelse(host_rank == 'family',new_host,family)) %>%
-  mutate(family = ifelse(host_rank == 'tribe',new_host,family)) %>%
-  rename(host=new_host,host_string = host_plant)
-head(fowler_formatted)
-
-#double check there's no genera in different families
-fowler_list = fowler_formatted %>% split(.$scientificName) 
-a = fowler_formatted %>% split(.$scientificName)
-df = a[[1]]
-check_these = which(fowler_formatted %>% split(.$scientificName) %>% map_lgl(function(df) length(unique(df$host))>1))
-generalists_i_think = which(check_these %>% purrr::map_dbl(function(i){
[…]

### 2023-03-07 scripts/eastern_western_bees.R
remove diller et al 2020 source for nonnative bees

@@ -22,6 +22,7 @@ my_dir2 = '/Volumes/Seagate/gbif_usa/'
 (my_paths1 = paste0(my_dir,list.files(my_dir1)))
 (my_paths2 = paste0(my_dir2,list.files(my_dir2)))
 my_paths = c(my_paths1, my_paths2)
+#
 
 #combine the paths together:
 bees = my_paths %>% map_dfr(function(path){

### 2023-03-07 scripts/make list of nonnatives.R

@@ -1,17 +1,18 @@
 rm(list=ls())
 library(tidyverse)
+library(vroom)
 
 # we need to exclude non-native species
-# first use list from USDA traits database (Diller et al 2020)
-nonnative = read_csv("/Users/colleen/Dropbox/My Mac (MacBook-Air.local)/Downloads/BeeGap_Taxonomy and General Traits.csv") %>% 
-  filter(`Native?` == "N") %>%  #N == 'no'
-  distinct(Genus,Species) %>%
-  mutate(scientificName = paste(Genus,Species), source = "Diller et al. 2020")
+# # first use list from USDA traits database (Diller et al 2020)
+# nonnative = read_csv("/Users/colleen/Dropbox/My Mac (MacBook-Air.local)/Downloads/BeeGap_Taxonomy and General Traits.csv") %>% 
+#   filter(`Native?` == "N") %>%  #N == 'no'
+#   distinct(Genus,Species) %>%
+#   mutate(scientificName = paste(Genus,Species), source = "Diller et al. 2020")
 
 
 # the list below is one I compiled from Russo 2016
 # https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5198217/
-more_nonnatives = c("Hylaeus variegates", "Hylaeus albonitens",
+nonnatives = c("Hylaeus variegates", "Hylaeus albonitens",
                     "Hylaeus hyalinatus", "Hylaeus leptocephalus","Hylaeus punctatus",
                     "Hylaeus strenuus","Andrena wilkella", "Halictus tectus", "Lasioglossum eleutherense",
                     "Lasioglossum leucozonium", "Lasioglossum zonulum", "Anthidium manicatum",
@@ -20,18 +21,76 @@ more_nonnatives = c("Hylaeus variegates", "Hylaeus albonitens",
                     "Megachile fullawayi", "Megachile lanata",	"Megachile rotundata", "Megachile sculpturalis", "Osmia caerulescens",
                     "Osmia cornifrons", "Osmia taurus", "Anthophora villosula", "Apis mellifera", "Ceratina cobaltina",
                     "Ceratina dallatorreana", "Euglossa dilemma", "Plebia frontalis", "Xylocopa appendiculata")
+#convert to data frame
+nonnatives_df = data.frame(scientificName = nonnatives ,source="Russo 2016") %>%
+  arrange(scientificName)
+
+
+#load chesshire et al, which has the name update info
[…]

### 2023-03-07 scripts/main analyses.R
adding scripts

@@ -20,15 +20,14 @@ library(gridExtra)
 # if we change the sample size how does this change?
 
 
-#load pollen data
-pollen_fidelity = read_csv("modeling_data/prop_host_pollen.csv")
-
 # load the globi data
 # this is a file with the globi records and plant names are updated (though not plant families)
 globi_r = read_csv('modeling_data/globi_occ_names_updated-19dec2022.csv')
 globi_r %>% filter(sourceTaxonRank=='variety') %>% distinct(scientificName)
 data.frame(globi_r %>% filter(is.na(sourceTaxonRank)) %>% distinct(scientificName))
 
+#this data file has bee and plant names updated, and only american bees
+globi = vroom('modeling_data/globi_american_native_bees.csv')
 
 #also load the fowler data
 diet_breadth = read_csv('modeling_data/bee_diet_breadth.csv')
@@ -136,46 +135,7 @@ american_sp_keep = c(american1, american2)[!c(american1,american2) %in% introduc
 globi_usa = globi_r %>% filter(scientificName %in% american_sp_keep) 
 
 
-# update the plant family names using world checklist for vascular plants
-# get_fams = unique(globi_usa$plant_genus) %>% future_map(function(sciname){
-#   output = search_wcvp(sciname, filters=c("families", "accepted"),limit = 1)
-#   output_df = tidy(output)
-#   
-#   if(nrow(output_df) !=0){
-#     return_df = data.frame(genus = sciname,family = output_df$family)
-#     
-#   }
-#   else{
-#     output = search_wcvp(sciname, filters=c("families"),limit = 1)
-#     df=tidy(output)
-#     
-#     if(nrow(df) !=0){
-#       new_output_df = df$synonymOf[[1]]
-#       return_df = data.frame(genus = sciname,family = new_output_df$family)
-#       
-#     }else{
-#       return_df = data.frame(genus=sciname,family = NA)
-#     }
-#   }
-# }) %>% bind_rows
-# saveRDS(get_fams,'modeling_data/globi_plant_fams.rds')
-get_fams = readRDS("modeling_data/globi_plant_fams.rds")
-
-#get_fams manually for things that are in the world vascular checklist (note some are tribes not genera)
-globi_r %>% filter(plant_genus =='Sarracenia') %>% select(referenceCitation)
[…]

### 2023-03-07 scripts/make plant phylogeny.R

@@ -1,0 +1,157 @@
+rm(list=ls())
+library(tidyverse)
+library(vroom)
+library(Taxonstand)
+library(hillR)
+library(purrr)
+library(vegan)
+#if vphylomaker not installed:
+#devtools::install_github("jinyizju/V.PhyloMaker")
+library(V.PhyloMaker)
+
+
+# load the filtered globi data
+# this is a file with the globi records and plant and bee names are updated 
+# bees are filtered to just be species that are native to the US
+globi = vroom('modeling_data/globi_american_native_bees.csv')
+
+#standardize according to the plant list, becaues that's what v.phylomaker uses
+plant_df = globi %>% distinct(plant_species,plant_genus,plant_family) %>% 
+  rename(genus=plant_genus,family=plant_family)
+test_me = plant_df$plant_species
+
+# #these are plants we already updated with TPL (takes a really long time to run)
+# check_names = read_csv('modeling_data/plant_list_name_update.csv')
+# 
+# #any plants not in the tpl data?
+# add_me = test_me[!test_me %in% check_names$Taxon] #yes - lots. let's add these
+# check_names2 = TPL(add_me) #takes forever to run, download output as csv
+# check_names_f = check_names %>% mutate(TPL.version = as.character(TPL.version)) %>%
+#   bind_rows(check_names2)
+
+# write_csv(check_names_f,'modeling_data/plant_list_name_update2.csv')
+check_names_f = read_csv('modeling_data/plant_list_name_update2.csv')
+
+#reformat the tpl data
+tpl_formatted = check_names_f %>% 
+  rename(plant_species = Taxon) %>% 
+  mutate(accepted_name = ifelse(!is.na(New.Species),paste(New.Genus,New.Species),New.Genus) ) %>%
+  mutate(accepted_name = ifelse(is.na(accepted_name), plant_species, accepted_name)) %>%
+  select(plant_species,accepted_name) %>%
+  filter(plant_species %in% plant_df$plant_species) %>%
+  rename(tpl_name = accepted_name)
+
+#make a dataframe with the family (from the wfo data) and the tpl genus and specise
+tpl_plants = tpl_formatted %>%
+  left_join(plant_df %>% select(plant_species,family)) %>%
+  mutate(genus = gsub(" .*","",tpl_name)) %>%
[…]

### 2023-03-07 scripts/name-alignment-chesshire.R

@@ -1,0 +1,36 @@
+rm(list=ls())
+library(tidyverse)
+library(readxl)
+
+chesshire2 = read_csv('modeling_data/Chesshire2023_nameAlignment.csv')
+chesshire = read_excel('modeling_data/Chesshire2023-appendixs3.xlsx') %>%
+  rename(finalName = "Final Name Selected",
+         providedName = 'Step 1: Original Scientific Name in Data Download',
+         secondRoundName = "Step 2: Chosen Name after comparing with DL List and/or first round of manual sleuthing",
+         reasonSecondRound = "Reason for Initial Name Change (or no change)",
+         JohnAscherReason = "John Notes and Final Decision")
+str(chesshire)
+unique(chesshire$reasonSecondRound)
+
+
+# let's look at records where the secondRoundName doesn't equal the final  name
+# these are most likely records where John Ascher disagreed with the DiscoverLife name
+# write_csv(chesshire %>%
+#   filter(finalName != secondRoundName) %>%
+#   select(providedName,secondRoundName,finalName,reasonSecondRound,JohnAscherReason),"modeling_data/look_at_me.csv")
+
+# let's upload same spreadsheet but with column saying whether the name alignment was informed by geography or not
+geo_informed = read_excel('modeling_data/Chesshire et al 2023 bees name alignment unsure.xlsx') %>%
+  filter(geographic_informed == 'yes')
+
+
+#filter original chesshire data to not include these species
+final = chesshire %>%
+  mutate(geo_informed = providedName %in% geo_informed$providedName) %>%
+  dplyr::select(providedName,finalName,geo_informed)
+
+#how many are informed by geography?
+nrow(geo_informed)
+
+  
+# write_csv(final,'modeling_data/Chesshire2023_nameAlignment-geoUpdate.csv')

### 2023-03-07 scripts/update bee names.R

@@ -1,0 +1,193 @@
+rm(list=ls())
+library(tidyverse)
+library(vroom)
+library(maps)
+library(sf)
+library(readxl)
+library(furrr)
+library(rgbif)
+#if vphylomaker not installed:
+#devtools::install_github("jinyizju/V.PhyloMaker")
+library(V.PhyloMaker)
+
+
+# load the globi data
+# this is a file with the globi records and plant names are updated 
+globi_r = read_csv('modeling_data/globi_occ_names_updated-19dec2022.csv')
+globi_r %>% filter(sourceTaxonRank=='variety') %>% distinct(scientificName)
+data.frame(globi_r %>% filter(is.na(sourceTaxonRank)) %>% distinct(scientificName))
+
+
+#also load the fowler data
+diet_breadth = read_csv('modeling_data/bee_diet_breadth.csv')
+specialists = diet_breadth[diet_breadth$diet_breadth=='specialist',]$scientificName
+fowler_formatted = read_csv('modeling_data/fowler_formatted-19dec2022.csv')
+
+#load list of non-native bee species
+nonnative = read_csv("modeling_data/nonnative_bees.csv")
+
+# Let's exclude interactions of non-US bees
+
+#first, load list of US bee species from discover life - the text file is from this link: 
+# https://www.discoverlife.org/nh/cl/counts/Apoidea_species.html
+usa_bees = read_table('modeling_data/Apoidea_species.txt',col_names=F) %>%
+  rename(genus = X1, epithet = X2) %>%
+  mutate(scientificName = paste(genus,epithet))
+
+#i don't think the discover life list is complete, 
+# so let's also check if these species are US bees by seeing if there
+# are at least 5 records in gbif
+
+
+#let's  remove subgenera in parentheses
+have_subgenera = data.frame(scientificName = globi_r$scientificName) %>% filter(grepl('(',scientificName,fixed=T)) %>% distinct(scientificName) %>% 
+  mutate(genus = gsub(" .*","",scientificName),
+                          epithet = sub(".*)",'', scientificName)) %>%
+  mutate(new_scientificName = paste0(genus,epithet))
+
+#update the globi data with the new species names
+globi_r_updated = globi_r %>% left_join(have_subgenera,by="scientificName") %>%
+  mutate(scientificName = ifelse(is.na(new_scientificName),scientificName,new_scientificName)) %>%
[…]

### 2023-03-07 scripts/eastern_western_bees.R
adding code

@@ -1,7 +1,6 @@
 rm(list=ls())
 library(tidyverse)
 library(vroom)
-library(rgbif)
 library(purrr)
 
 
@@ -19,61 +18,276 @@ library(purrr)
 #load the idigbio and gbif data
 my_dir1 = '/Volumes/Seagate/idigbio_usa/'
 my_dir2 = '/Volumes/Seagate/gbif_usa/'
-(my_paths1 = paste0(my_dir,list.files(my_dir1)))
+(my_paths1 = paste0(my_dir1,list.files(my_dir1)))
 (my_paths2 = paste0(my_dir2,list.files(my_dir2)))
 my_paths = c(my_paths1, my_paths2)
 #
 
-#combine the paths together:
-bees = my_paths %>% map_dfr(function(path){
+#load the idigbio data and rename columns
+bees_idig = my_paths1 %>% map_dfr(function(path){
   vroom(path) %>% 
-    mutate("dwc:recordNumber" = as.character("dwc:recordNumber")) %>%
-    rename(scientificName = `dwc:scientificName`,
-           coordinateUncertaintyInMeters = `dwc:coordinateUncertaintyInMeters`,
-           stateProvince = `dwc:stateProvince`)
-  
-  })
+    mutate("dwc:recordNumber" = as.character("dwc:recordNumber")) 
+  })%>%
+  rename(scientificName = `dwc:scientificName`,
+         coordinateUncertaintyInMeters = `dwc:coordinateUncertaintyInMeters`,
+         stateProvince = `dwc:stateProvince`,
+         coords = `idigbio:geoPoint`,
+         genus = `dwc:genus`,
+         specificEpithet = `dwc:specificEpithet`) %>%
+  mutate(decimalLongitude = as.numeric(sub("}.*","",sub(".*: ","",coords)))) %>% #format longitude
+  mutate(decimalLatitude = as.numeric(sub(",.*","",sub(".*at\": ","",coords)))) %>%#format latitude
+  filter(!'dwc:basisOfRecord' %in% c( 'fossilspecimen','machineobservation'))
 
-test=vroom(my_paths2[[1]])
+
+#since there are so many columns for the gbif data
+#we'll just select the ones we want
+my_paths3 = my_paths2[!grepl("apid",my_paths2)]#since apidae is huge and takes a long time to load let's try doing it separately from the other files
+apidae_path = my_paths2[grepl("apid",my_paths2)]
+bees_gbif2 = my_paths3 %>% 
+  map_dfr(function(path){
[…]

### 2023-03-07 scripts/Chesshire2023-beeArea.R
code for calculating area of convex hull

@@ -1,0 +1,104 @@
+#get bee area
+rm(list=ls())
+library(tidyverse)
+library(vroom)
+library(purrr)
+library(sf)
+library(sp)
+library(grDevices)
+library(BBmisc)
+
+#
+my_dir = '/Volumes/Seagate/Chesshire2023'
+list.files(my_dir)
+
+#cleaned occurence records from chesshire et al
+bees = fread(paste0(my_dir,"contiguousRecords_high_Only.csv"))
+
+#get distinct updated names
+bee_names = bees %>% distinct(scientificName, finalName) %>%
+  rename(providedName = scientificName)
+# write_csv(bee_names,paste0(my_dir,'Chesshire2023_nameAlignment.csv'))
+# write_csv(bee_names,'modeling_data/Chesshire2023_nameAlignment.csv')
+
+# get rid of non-natives 
+nonnative=read_csv('modeling_data/nonnative_bees.csv') #need to realign using same naming conventions
+bees_filtered = bees  %>% # only bees we have aligned names for
+  filter(!finalName %in% nonnative$scientificName)  #get rid of non-natives
+n_distinct(bees_filtered$finalName)
+bees_filtered_distinct = bees_filtered %>%
+  distinct(finalName,finalLatitude,finalLongitude) 
+nrow(bees_filtered)#weird, shouldn't these be the same? I thought unique lat-longs were taken for each bee
+nrow(bees_filtered_distinct)
+
+#get rid of bees with fewer than 4 records
+bees_4ormore = bees_filtered_distinct %>% group_by(finalName) %>%
+  summarize(n=n()) %>%
+  filter(n>=4)
+
+my_list = bees_filtered_distinct %>% filter(finalName %in% bees_4ormore$finalName) %>%
+  split(.$finalName)
+sp = my_list[[1]]
+
+bee_area = bees_filtered_distinct %>% 
+  filter(finalName %in% bees_4ormore$finalName) %>%
+  split(.$finalName) %>%
+  purrr::map_dfr(function(sp){
+    ch <- chull(sp$finalLongitude,sp$finalLatitude)
+    coords = sp %>% select(finalLongitude,finalLatitude) %>% slice(ch,ch[1])
+    pols <- SpatialPolygons(list(Polygons(list(Polygon(coords)), 1)), proj4string = CRS("+proj=longlat +datum=WGS84"))
[…]

### 2023-03-07 scripts/update bee names.R
change to chesshire et al method of name alignment

@@ -6,6 +6,9 @@ library(sf)
 library(readxl)
 library(furrr)
 library(rgbif)
+library(sp)
+library(tigris)
+
 #if vphylomaker not installed:
 #devtools::install_github("jinyizju/V.PhyloMaker")
 library(V.PhyloMaker)
@@ -13,10 +16,11 @@ library(V.PhyloMaker)
 
 # load the globi data
 # this is a file with the globi records and plant names are updated 
-globi_r = read_csv('modeling_data/globi_occ_names_updated-19dec2022.csv')
+globi_r = read_csv('modeling_data/globi_occ_names_updated-19dec2022.csv') %>%
+  mutate(unique_id = 1:nrow(.)) 
+
 globi_r %>% filter(sourceTaxonRank=='variety') %>% distinct(scientificName)
 data.frame(globi_r %>% filter(is.na(sourceTaxonRank)) %>% distinct(scientificName))
-
 
 #also load the fowler data
 diet_breadth = read_csv('modeling_data/bee_diet_breadth.csv')
@@ -26,6 +30,159 @@ fowler_formatted = read_csv('modeling_data/fowler_formatted-19dec2022.csv')
 #load list of non-native bee species
 nonnative = read_csv("modeling_data/nonnative_bees.csv")
 
+#chesshire et al data for aligning names
+chesshire = read_csv('modeling_data/Chesshire2023_nameAlignment-geoUpdate.csv') %>% 
+  filter(providedName != "Megachile concinna")
+#exclude Megachile conccina --> native species, name change from Chesshire since
+#  according to Russo 2016, M coccina is actually non-native  
+
+# format names:
+#let's  remove subgenera in parentheses
+have_subgenera = data.frame(scientificName = globi_r$scientificName) %>% filter(grepl('(',scientificName,fixed=T)) %>% distinct(scientificName) %>% 
+  mutate(genus = gsub(" .*","",scientificName),
+         epithet = sub(".*)",'', scientificName)) %>%
+  mutate(new_scientificName = paste0(genus,epithet))
+
+#update the globi data with the new species names
+globi_r_updated = globi_r %>% left_join(have_subgenera,by="scientificName") %>%
+  mutate(scientificName = ifelse(is.na(new_scientificName),scientificName,new_scientificName)) %>%
+  mutate(scientificName = ifelse(grepl("Apis mellifera",scientificName),"Apis mellifera", scientificName)) #remove the weirdly formatted Apis mellifera
+
+#get the list of species
[…]

### 2023-03-07 scripts/update bee names.R
changing to chesshire et al name alignment methods

@@ -1,17 +1,11 @@
 rm(list=ls())
 library(tidyverse)
 library(vroom)
-library(maps)
 library(sf)
 library(readxl)
-library(furrr)
-library(rgbif)
 library(sp)
 library(tigris)
 
-#if vphylomaker not installed:
-#devtools::install_github("jinyizju/V.PhyloMaker")
-library(V.PhyloMaker)
 
 
 # load the globi data

### 2023-03-07 scripts/format_fowler_hosts.R
changing to chesshire et al name alignment methods

@@ -19,6 +19,14 @@ not_in_wfo = data.frame(fowler = c('Salazaria'),
                         family = c("Lamiaceae"),
                         source = c('col'))
 
+#chesshire et al data for aligning names
+chesshire1 = read_csv('modeling_data/Chesshire2023_nameAlignment-geoUpdate.csv') 
+#for final names that also aren't in provided name: add them
+add_chesshire = data.frame(providedName = chesshire1$finalName[!chesshire1$finalName %in% chesshire1$providedName]) %>%
+  mutate(finalName = providedName)
+
+chesshire = bind_rows(chesshire1,add_chesshire)
+
 # fowler considers bees to be specialists if they use pollen from genera in two different families
 #and we don't. we're going to remove these bees from the fowler list (and won't consider them discrepancies)
 generalists_fowler = c("Andrena candidiformis", "Anthidium mormonum", "Dufourea cuprea",
@@ -191,9 +199,31 @@ fowler_formatted = hosts_long %>% rename(old_plant = new_host) %>%
 
 nrow(fowler_formatted) == nrow(hosts_long)
 
-#next we need to update the bee names
-#do this using name alignment template
+#next we need to update the bee names using chesshire's methods
+fowler_names_toAlign = data.frame(providedName = unique(fowler_formatted$scientificName))
 
+slist = fowler_names_toAlign$providedName
+slist[!slist %in% chesshire$providedName]
+
+folwer_updated = fowler_names_toAlign %>%
+  left_join(chesshire) %>%
+  mutate(finalName = ifelse(is.na(finalName),providedName,finalName)) %>% #for bees not in chesshire et al, just use the provided name
+  dplyr::select(-geo_informed)
+
+#now update teh fowler dataframes
+fowler_formatted2=fowler_formatted %>% 
+  rename(old_bee_name = scientificName) %>%
+  left_join(folwer_updated %>% rename(old_bee_name = providedName,scientificName = finalName))
+diet_breadth2 = diet_breadth %>% 
+  rename(old_bee_name = scientificName) %>%
+  left_join(folwer_updated %>% rename(old_bee_name = providedName,scientificName = finalName))
+
+
+# write_csv(fowler_formatted2,'modeling_data/fowler_formatted-7march2023.csv')
+# write_csv(diet_breadth2,'modeling_data/bee_diet_breadth-7march2023.csv')
+
+##old method of bee name alignment
+#do this using name alignment template
[…]

### 2023-03-08 modeling_data/GLOBI_bee_update_clean.Rmd
commenting out write_csv

@@
 output:         html_document
-participants:   N. Bachelder, A. Chen, Z. Fang, M. Rapaport, S. Solomon
+participants:   N. Bachelder, A. Chen, Z. Fang, M. Rapaport, S. Solomon, C. Smith
 date:           Feb 18, 2021
@@
 ```{r}
-readr::write_csv(interactions, "interactions-14dec2022.csv")
+# readr::write_csv(interactions, "interactions-14dec2022.csv")
 ```

### 2023-03-08 modeling_data/GLOBI_bee_update_clean.Rmd
more reformatting

@@
 # Combine and drop duplicates
-interactions = rbind(bee_source_df, plant_source_df) %>% unique()
+interactions_b = rbind(bee_source_df, plant_source_df) %>% unique()
+```
+
+## Reformat column names
+```{r}
+for (c in colnames(interactions_b)) {
+  interactions_b[[c]] <-  tolower(interactions_b[[c]])
+}
+
+capitalize = function(a_string){
+
+  beginning = toupper(substr(a_string,1,1))
+  ending = substr(a_string,2,nchar(a_string))
+
+  return(paste0(beginning,ending))
+
+}
+
+```
+
+## Remove plants not id'd to at least genus level, and bees that are not id'd to species
+## Change old family names for Asteraceae, Lamiaceae and Apiaceae
+```{r}
+#reformat the interactions data:
+interactions_reformat = interactions_b %>% 
+  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
+         bee_family = capitalize(sourceTaxonFamilyName),
+         plant_family = capitalize(targetTaxonFamilyName), 
+         plant_genus = capitalize(targetTaxonGenusName),
+         plant_species = capitalize(targetTaxonName)) %>%
+  filter(scientificName != '' & plant_family !='' & plant_genus != "") %>%
+  filter(targetTaxonRank %in% c('subspecies','species',"genus",'subgenus','variety')) %>%
+  filter(sourceTaxonRank %in% c("subspecies",'variety','species')) %>%
+  mutate(plant_family = ifelse(plant_family=='Compositae','Asteraceae',plant_family)) %>%
+  mutate(plant_family = ifelse(plant_family=='Umbelliferae','Apiaceae',plant_family)) %>%
+  mutate(plant_family = ifelse(plant_family=='Labiatae','Lamiaceae',plant_family)) 
 ```
@@
 ```{r}
-# readr::write_csv(interactions, "interactions-14dec2022.csv")
+# readr::write_csv(interactions_reformat, "interactions-14dec2022.csv")
 ```

### 2023-03-08 scripts/update bee names.R
changing order of workflow

@@ -9,8 +9,10 @@ library(tigris)
 
 
 # load the globi data
-# this is a file with the globi records and plant names are updated 
-globi_r = read_csv('modeling_data/globi_occ_names_updated-19dec2022.csv') %>%
+
+# this is a file with the globi records filtered to be just bees 
+# without plant names updated 
+globi_r = vroom("modeling_data/interactions-14dec2022.csv")%>%
   mutate(unique_id = 1:nrow(.)) 
 
 globi_r %>% filter(sourceTaxonRank=='variety') %>% distinct(scientificName)
@@ -76,7 +78,7 @@ globi_align1 = globi_provided %>%
   filter(!is.na(finalName)) %>%
   bind_rows(only_in_final)
 
-#there are also bees where the provided names are mispeleld or incorrect but the 
+#there are also bees where the provided names are misspelled or incorrect but the 
 # alignment isn't in the chesshire data
 # add these
 alignment_toAdd = data.frame(providedName = c("Melissodes lupina","Melissodes bimaculata","Melissodes rustica"),
@@ -86,7 +88,7 @@ globi_align2 =  globi_provided %>%
   left_join(alignment_toAdd) %>%
   filter(!is.na(finalName)) 
 
-# second do alignments that are informed by geography
+# next do alignments that are informed by geography
 chesshire_geo = chesshire %>% filter(geo_informed)
 globi_provided %>%
   filter(providedName %in% chesshire_geo$providedName) #only 5
@@ -133,7 +135,7 @@ rm_uids %>%
 #how many records are removed?
 nrow(rm_uids)
 
-#update globi daata again
+#update globi data again
 globi_r_updated2 = globi_r_updated %>%
   filter(!unique_id %in% rm_uids$unique_id)
 #make third name alignment df for name updates informed by geography
@@ -162,12 +164,12 @@ old_globi = read_csv('modeling_data/globi_american_native_bees.csv')
   group_by(old_bee_name, scientificName) %>%
   summarize(n=n()) %>%
   arrange(desc(n)))
-check_me_bees %>% filter(old_bee_name != scientificName) #this species is alaskan...
+check_me_bees %>% filter(old_bee_name != scientificName) #this species is alaskan
 
 #
 
 
-# write_csv(globi_names1,'modeling_data/globi_american_native_bees_7march2023.csv')
+ # write_csv(globi_names1,'modeling_data/globi_american_native_bees_7march2023.csv')
 
 
 

### 2023-03-08 scripts/make plant phylogeny.R
changing dataset

@@ -13,85 +13,125 @@ library(V.PhyloMaker)
 # load the filtered globi data
 # this is a file with the globi records and plant and bee names are updated 
 # bees are filtered to just be species that are native to the US
-globi = vroom('modeling_data/globi_american_native_bees.csv')
+globi = vroom('modeling_data/globi_american_native_bees_7march2023.csv')
 
-#standardize according to the plant list, becaues that's what v.phylomaker uses
+#standardize according to the plant list, because that's what v.phylomaker uses
 plant_df = globi %>% distinct(plant_species,plant_genus,plant_family) %>% 
   rename(genus=plant_genus,family=plant_family)
 test_me = plant_df$plant_species
 
-# #these are plants we already updated with TPL (takes a really long time to run)
-# check_names = read_csv('modeling_data/plant_list_name_update.csv')
-# 
-# #any plants not in the tpl data?
-# add_me = test_me[!test_me %in% check_names$Taxon] #yes - lots. let's add these
-# check_names2 = TPL(add_me) #takes forever to run, download output as csv
-# check_names_f = check_names %>% mutate(TPL.version = as.character(TPL.version)) %>%
-#   bind_rows(check_names2)
+#we'll get families from wfo
+#and also do genus updates with those, since tpl doesn't have a lot of those
+genus_updates = read_csv('modeling_data/wfo_genus_updates.csv')
+wfo_fams = read_csv('modeling_data/plant_fam_info.csv')
 
-# write_csv(check_names_f,'modeling_data/plant_list_name_update2.csv')
-check_names_f = read_csv('modeling_data/plant_list_name_update2.csv')
+# # ##uncomment me to check you have all the names
+# # # #these are plants we already updated with TPL (takes a really long time to run)
+# check_names_f = read_csv('modeling_data/plant_list_name_update2.csv')
+# # # #any plants not in the tpl data?
+# add_me = test_me[!test_me %in% check_names_f$Taxon] #yes - lots. let's add these
+# check_names2 = TPL(add_me) #takes forever to run, download output as csv
+# check_names = check_names_f %>% mutate(TPL.version = as.character(TPL.version)) %>%
[…]

### 2023-03-09 scripts/Chesshire2023-beeArea.R
combining data

@@ -7,9 +7,12 @@ library(sf)
 library(sp)
 library(grDevices)
 library(BBmisc)
+library(data.table)
+library(readxl)
+library(lubridate)
 
 #
-my_dir = '/Volumes/Seagate/Chesshire2023'
+my_dir = '/Volumes/Seagate/Chesshire2023/'
 list.files(my_dir)
 
 #cleaned occurence records from chesshire et al
@@ -22,13 +25,14 @@ bee_names = bees %>% distinct(scientificName, finalName) %>%
 # write_csv(bee_names,'modeling_data/Chesshire2023_nameAlignment.csv')
 
 # get rid of non-natives 
-nonnative=read_csv('modeling_data/nonnative_bees.csv') #need to realign using same naming conventions
+nonnative=read_excel('modeling_data/nonnative_bees.xlsx') #need to realign using same naming conventions
+
 bees_filtered = bees  %>% # only bees we have aligned names for
   filter(!finalName %in% nonnative$scientificName)  #get rid of non-natives
 n_distinct(bees_filtered$finalName)
 bees_filtered_distinct = bees_filtered %>%
-  distinct(finalName,finalLatitude,finalLongitude) 
-nrow(bees_filtered)#weird, shouldn't these be the same? I thought unique lat-longs were taken for each bee
+  distinct(finalName,finalLatitude,finalLongitude,eventDate) 
+nrow(bees_filtered)#weird, shouldn't these be the same? I thought unique lat-longs were taken for each bee (maybe per collection event?)
 nrow(bees_filtered_distinct)
 
 #get rid of bees with fewer than 4 records
@@ -69,20 +73,14 @@ bee_area = bees_filtered_distinct %>%
 
 ##
 #double check the convex hull code by plotting
-plot(coords,pch=16,col=adjustcolor("black",.8))
-lines(coords[ch,])
-
-
-
-##
-# Data
-sp = my_list[[10]]
-
+a=bees_filtered_distinct %>% 
+  filter(finalName %in% bees_4ormore$finalName) %>%
+  split(.$finalName)
+sp = a[[2]]
 ch <- chull(sp$finalLongitude,sp$finalLatitude)
 coords = sp %>% select(finalLongitude,finalLatitude) %>% slice(ch,ch[1])
+# plot(coords,pch=16,col=adjustcolor("black",.8))
[…]

### 2023-03-09 scripts/finalDataset.R

@@ -1,0 +1,31 @@
+rm(list=ls())
+library(tidyverse)
+
+bee_phy = read_csv('modeling_data/bee_phylogenetic_data.csv')
+plant_phy = read_csv("modeling_data/globi_phyloDiv.csv") %>%
+  mutate(bee_genus = sub(" .*","",scientificName))
+geo = read_csv('modeling_data/chesshire2023_beeArea.csv')
+diet_breadth = read_csv('modeling_data/bee_diet_breadth-7march2023.csv')
+
+
+#plant_phy has all the bees in globi
+#combine with bee_phy and geo
+#add diet breadth info
+data = plant_phy %>%
+  left_join(bee_phy, by = 'bee_genus') %>%
+  left_join(geo)  %>% left_join(diet_breadth %>% 
+  distinct(scientificName,diet_breadth,diet_breadth_detailed)) %>%
+  mutate(diet_breadth = ifelse(is.na(diet_breadth),'generalist',diet_breadth),
+         diet_breadth_detailed = ifelse(is.na(diet_breadth_detailed),'generalist',diet_breadth_detailed)) %>%
+  select(scientificName,bee_genus,bee_family,diet_breadth,everything())
+
+
+#variables from the data we care about:
+# phylo_rich,phylo_simp,n_chesshire,area_ha,mean_doy,eigen1,eigen2, mean_lat,mean_long
+
+View(data %>% 
+       select(scientificName,phylo_rich,phylo_simp,n_chesshire,area_ha,med_doy,flight_season,med_lat,med_long,eigen1,eigen2))
+
+#let's see if randomForest works with NA values for area
+library(randomForest)
+rf = randomForest(as.factor(diet_breadth) ~ phylo_rich + area_ha,importance = T,data=data)

### 2023-03-09 scripts/update bee names.R

@@ -169,7 +169,7 @@ check_me_bees %>% filter(old_bee_name != scientificName) #this species is alaska
 #
 
 
- # write_csv(globi_names1,'modeling_data/globi_american_native_bees_7march2023.csv')
+# write_csv(globi_names1,'modeling_data/globi_american_native_bees_7march2023.csv')
 
 
 

### 2023-03-09 scripts/update plant names.R
combining data

@@ -5,38 +5,11 @@ library(tidyverse)
 
 #load the data
 options(max.print=1000000)
-interactions_b <- vroom('modeling_data/interactions-14dec2022.csv')
+interactions_b <- vroom('modeling_data/interactions-14dec2022.csv') #note this is non-filtered data (includes non-usa & non-native bees)
 fowler <- read_csv("modeling_data/fowler_hostplants.csv") 
 
-for (c in colnames(interactions_b)) {
-  interactions_b[[c]] <-  tolower(interactions_b[[c]])
-}
 
-
-capitalize = function(a_string){
-  
-  beginning = toupper(substr(a_string,1,1))
-  ending = substr(a_string,2,nchar(a_string))
-  
-  return(paste0(beginning,ending))
-  
-}
-
-#reformat the interactions data:
-interactions_reformat = interactions_b %>% 
-  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
-         bee_family = capitalize(sourceTaxonFamilyName),
-         plant_family = capitalize(targetTaxonFamilyName), 
-         plant_genus = capitalize(targetTaxonGenusName),
-         plant_species = capitalize(targetTaxonName)) %>%
-  filter(scientificName != '' & plant_family !='' & plant_genus != "") %>%
-  filter(targetTaxonRank %in% c('subspecies','species',"genus",'subgenus','variety')) %>%
-  filter(sourceTaxonRank %in% c("subspecies",'variety','species')) %>%
-  mutate(plant_family = ifelse(plant_family=='Compositae','Asteraceae',plant_family)) %>%
-  mutate(plant_family = ifelse(plant_family=='Umbelliferae','Apiaceae',plant_family)) %>%
-  mutate(plant_family = ifelse(plant_family=='Labiatae','Lamiaceae',plant_family)) 
-
-plant_df = interactions_reformat %>%
+plant_df = interactions_b %>%
   distinct(plant_species)
 test_me = plant_df$plant_species
 
@@ -56,11 +29,8 @@ not_in_wfo = data.frame(globi_name = c("Spiranthes arcisepala","Lisianthus skinn
[…]

### 2023-03-09 scripts/make bee phylogeny.R
adding script

@@ -1,0 +1,106 @@
+rm(list=ls())
+library(tidyverse)
+library("ape")
+library("phytools")
+library('picante')
+
+##if U.phylomaker is not installed:
+# devtools::install_github("jinyizju/U.PhyloMaker")
+library("U.PhyloMaker")
+
+#### Read in phylogenetic tree
+mytree <- read.tree("modeling_data/12862_2013_2375_MOESM1_ESM.txt",comment.char = "#", keep.multi = TRUE, tree.names=c("treeA","treeB","treeC","treeD","treeE","treeF","treeG","treeH","treeI","treeJ"))
+
+###read in the globi bee data
+globi = read_csv("modeling_data/globi_allNamesUpdated.csv") %>%
+  mutate(bee_genus = sub(" .*","",scientificName))
+globi_genera = unique(globi$bee_genus)
+globi_species2 = unique(globi$scientificName)
+globi_species =sub(" ","_",globi_species2)
+
+# abritrarily picked the first tree of a list of 10 to work with
+workingtree <- mytree[[1]] 
+
+
+#make data frame with the bees in the tree and their genus
+tree_bees = data.frame(bee = workingtree$tip.label) %>%
+  mutate(bee_genus = sub("_.*",'',bee))
+
+#any of the globi bee genera not in the tree?
+(genera_out = globi_genera[!globi_genera %in% tree_bees$bee_genus] ) #yup
+bees_in_tree = globi_species[globi_species %in% tree_bees$bee]
+
+set.seed(7211)
+#for bee genera with species in the tree randomly pick one of each species
+bees_in_tree_sampled = globi %>% filter(scientificName %in% sub("_"," ",bees_in_tree)) %>%
+  split(.$bee_genus) %>% map_dfr(function(df){
+    sampled_index = sample(nrow(df),1) #get a random row index
+    output = df[sampled_index,] %>% select(scientificName) %>% 
+      rename(bee = scientificName) %>% mutate(bee_genus = sub(" .*","",bee))
+    return(output)
+  })
+
+#for bee genera that are on the tree (but different species)
+#randomly pick a species that's on the tree
+bee_gen_in_tree = globi_genera[!globi_genera %in% c(genera_out,bees_in_tree_sampled$bee_genus)]
[…]

### 2023-03-09 scripts/finalDataset.R
adding data

@@ -20,9 +20,11 @@ data = plant_phy %>%
   select(scientificName,bee_genus,bee_family,diet_breadth,everything())
 
 
+#write final dataset to csv
+# write_csv(data,"modeling_data/globi_speciesLevelFinal.csv")
 #variables from the data we care about:
 # phylo_rich,phylo_simp,n_chesshire,area_ha,mean_doy,eigen1,eigen2, mean_lat,mean_long
-
+View(data)
 View(data %>% 
        select(scientificName,phylo_rich,phylo_simp,n_chesshire,area_ha,med_doy,flight_season,med_lat,med_long,eigen1,eigen2))
 

### 2023-03-09 scripts/finalDataset.R
changing column order

@@ -12,12 +12,12 @@ diet_breadth = read_csv('modeling_data/bee_diet_breadth-7march2023.csv')
 #combine with bee_phy and geo
 #add diet breadth info
 data = plant_phy %>%
-  left_join(bee_phy, by = 'bee_genus') %>%
-  left_join(geo)  %>% left_join(diet_breadth %>% 
+  left_join(geo) %>%
+  left_join(bee_phy)  %>% left_join(diet_breadth %>% 
   distinct(scientificName,diet_breadth,diet_breadth_detailed)) %>%
   mutate(diet_breadth = ifelse(is.na(diet_breadth),'generalist',diet_breadth),
          diet_breadth_detailed = ifelse(is.na(diet_breadth_detailed),'generalist',diet_breadth_detailed)) %>%
-  select(scientificName,bee_genus,bee_family,diet_breadth,everything())
+  select(scientificName,bee_genus,bee_family,diet_breadth,diet_breadth_detailed,everything())
 
 
 #write final dataset to csv

### 2023-03-09 scripts/finalDataset.R
test commit

@@ -27,7 +27,7 @@ data = plant_phy %>%
 View(data)
 View(data %>% 
        select(scientificName,phylo_rich,phylo_simp,n_chesshire,area_ha,med_doy,flight_season,med_lat,med_long,eigen1,eigen2))
-
+#
 #let's see if randomForest works with NA values for area
 library(randomForest)
 rf = randomForest(as.factor(diet_breadth) ~ phylo_rich + area_ha,importance = T,data=data)

### 2023-03-09 scripts/make bee phylogeny.R
changing to genus-level tree, rooted, ultrametric

@@ -9,78 +9,54 @@ library('picante')
 library("U.PhyloMaker")
 
 #### Read in phylogenetic tree
-mytree <- read.tree("modeling_data/12862_2013_2375_MOESM1_ESM.txt",comment.char = "#", keep.multi = TRUE, tree.names=c("treeA","treeB","treeC","treeD","treeE","treeF","treeG","treeH","treeI","treeJ"))
+# mytree <- read.tree("modeling_data/12862_2013_2375_MOESM1_ESM.txt",comment.char = "#", keep.multi = TRUE, tree.names=c("treeA","treeB","treeC","treeD","treeE","treeF","treeG","treeH","treeI","treeJ"))
+mytree <- read.tree('modeling_data/12862_2013_2375_MOESM3_ESM.txt')
 
-###read in the globi bee data
+### read in the globi bee data
 globi = read_csv("modeling_data/globi_allNamesUpdated.csv") %>%
   mutate(bee_genus = sub(" .*","",scientificName))
 globi_genera = unique(globi$bee_genus)
 globi_species2 = unique(globi$scientificName)
 globi_species =sub(" ","_",globi_species2)
 
-# abritrarily picked the first tree of a list of 10 to work with
-workingtree <- mytree[[1]] 
+# abritrarily picked the a tree from the list of 10 to work with
+which_index = 1
+bee_tree <- mytree[[which_index]] 
+
+###root with apoid wasp outgroup
+workingtree=root(bee_tree,outgroup="Tachysphex")
+workingtree=as.phylo(workingtree)
+# 
+##Make ultrametric
+workingtree=chronos(workingtree)
+# is.ultrametric(bee_tree)
+
+#remove bee genera that are not in globi
+drop_me = bee_tree$tip.label[!bee_tree$tip.label %in% globi_genera]
+pruned_tree = drop.tip(workingtree, drop_me)
 
 
-#make data frame with the bees in the tree and their genus
-tree_bees = data.frame(bee = workingtree$tip.label) %>%
-  mutate(bee_genus = sub("_.*",'',bee))
+# add genera not in Hedtke
+(genera_out = globi_genera[!globi_genera %in% workingtree$tip.label])#look at them
+species_list = globi_genera
+genus_list = data.frame(globi %>%
[…]

### 2023-03-10 scripts/Chesshire2023-beeArea.R
adding more data

@@ -16,7 +16,8 @@ my_dir = '/Volumes/Seagate/Chesshire2023/'
 list.files(my_dir)
 
 #cleaned occurence records from chesshire et al
-bees = fread(paste0(my_dir,"contiguousRecords_high_Only.csv"))
+#to be used for usa species list:
+bees = fread(paste0(my_dir,"contiguousRecords_high_Only.csv")) #this contains only contiguous USA records
 
 #get distinct updated names
 bee_names = bees %>% distinct(scientificName, finalName) %>%
@@ -24,43 +25,137 @@ bee_names = bees %>% distinct(scientificName, finalName) %>%
 # write_csv(bee_names,paste0(my_dir,'Chesshire2023_nameAlignment.csv'))
 # write_csv(bee_names,'modeling_data/Chesshire2023_nameAlignment.csv')
 
-# get rid of non-natives 
-nonnative=read_excel('modeling_data/nonnative_bees.xlsx') #need to realign using same naming conventions
 
-bees_filtered = bees  %>% # only bees we have aligned names for
-  filter(!finalName %in% nonnative$scientificName)  #get rid of non-natives
+#rm bees since its so big
+rm(bees)
+#upload data with records for all of NA
+#to be used for estimating extent of occurrence
+allNA = fread(paste0(my_dir,"NorAmer_highQual_only_ALLfamilies.csv"))#this contains records records from Mexico, Canada, Alaska 
+
+# get rid of non-natives 
+nonnative=read_excel('modeling_data/nonnative_bees.xlsx') 
+
+#filter North American dataset to just be bees in bee_names
+#and just be native species
+#also add bee average collection date
+#dates in this df have two different formats that we need to account for
+bees_filtered = allNA  %>% # only bees we have aligned names for
+  filter(!finalName %in% nonnative$scientificName) %>%  #get rid of non-natives
+  filter(finalName %in% bee_names$finalName) %>%
+  mutate(fdate = ifelse(grepl('/',eventDate),as.character(mdy(eventDate)),as.character(as.Date(eventDate)))) %>%
+  mutate(doy = yday(as.character(fdate)))
+
 n_distinct(bees_filtered$finalName)
+bees_filtered[1498290,]
+bees_filtered %>%
+  filter(is.na(fdate)) %>%
+  distinct(eventDate) %>% #double check that the only thing that isn't getting reformatted is NAs
[…]

### 2023-12-21 modeling_data/old_files/Full Bredth v Specialization Study.Rmd
moving files

--- 2022-06-06 Full Bredth v Specialization Study.Rmd
@@
 # Remove cases where bee rarity changes across area # DONE
-# Change common catagory to 'not-rare' (assumption: fowler is exhustive) # DONE
+# Change common catagory to 'not-rare' (assumption: fowler is exhaustive) # DONE
 # location can be tossed due to data limitation, use broad scale east, west, central together.
@@
 library(ggplot2)
-compare_df <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/compare_df.csv') %>% dplyr::select(-X)
+compare_df <- read.csv('modeling_data/compare_df.csv')
 

### 2024-02-21 modeling_data/GLOBI_bee_update_clean.Rmd
adding script that got deleted

@@
 ```{r}
+#some of these are not vascular plants or not id'd to genus (e.g Heliantheae)
+rm_me = c("Coleoptera","White","Picconiella","Heliantheae","Eupatorieae","Eremocarpus") #Eremocarpus is now split into 3 different genera
+
 #reformat the interactions data:
@@
   mutate(plant_family = ifelse(plant_family=='Umbelliferae','Apiaceae',plant_family)) %>%
-  mutate(plant_family = ifelse(plant_family=='Labiatae','Lamiaceae',plant_family)) 
+  mutate(plant_family = ifelse(plant_family=='Labiatae','Lamiaceae',plant_family)) %>% filter(!plant_species %in% rm_me)
+
 ```
